---
title: "Mycoplasma_vs_Synechocystis_vs_E.coli"
author: "Mikhail Orlov"
date: '6 ноября 2017 г '
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clearing workspace, setting working directory and loading R packages
```{r, message=FALSE}
rm(list=ls())

#setwd('/home/mikhail/Documents/Script_2016_all/PCA/')

library(R.matlab)
library(seqinr)
library(factoextra)
#library(data.table)
library(caret)
#library(doMC)
library(Biostrings)
library(ape)
library(reldna)
library(fastcluster)
library(dendextend)
library(ggplot2)
library(ggseqlogo)


```
Reading in E. coli K12 genome (GenBank Accesion U00096.2), creating reverse complement genome and tranforming genome into character form 
```{r Reading in E. coli genome}
e.coli_U00096.2<-paste0(read.GenBank(access.nb = 'U00096.2', as.character = T)$U00096.2, collapse = '')
reverseComplement_e.coli_U00096.2<-as.character(reverseComplement(DNAString(e.coli_U00096.2)))

e.coli_U00096.2_char<-unlist(strsplit(e.coli_U00096.2, ''))
reverseComplement_e.coli_U00096.2_char<-unlist(strsplit(reverseComplement_e.coli_U00096.2, ''))

```

Reading in Synechocystis sp. PCC6803 complete, creating reverse complement genome and tranforming genome into character form 
```{r Reading in Synechocystis sp. PCC6803 complete genome, in supplementary noted genbank acession NC_000911.1, warning=FALSE, message=FALSE, cache=TRUE}
synechocystis_NC_000911.1 <- toupper(paste0(read.GenBank(access.nb = 'NC_000911.1', as.character = T)$NC_000911.1, collapse = ''))
reverseComplement_synechocystis_NC_000911.1 <- as.character(reverseComplement(DNAString(synechocystis_NC_000911.1)))

synechocystis_NC_000911.1_char<-unlist(strsplit(synechocystis_NC_000911.1, ''))
reverseComplement_synechocystis_NC_000911.1_char<-unlist(strsplit(reverseComplement_synechocystis_NC_000911.1, ''))

```

Reading in M. gallisepticum strain S6 genome (CP006916.3 assembly), creating reverse complement genome and tranforming genome into character form 
```{r Reading in M. gallisepticum strain S6 genome (CP006916.3 assembly), warning=FALSE, message=FALSE, cache=TRUE}

mycoplasma_CP006916.3<- toupper(paste0(read.GenBank(access.nb = 'CP006916.3', as.character = T)$CP006916.3, collapse = ''))
reverseComplement_mycoplasms_CP006916.3<- as.character(reverseComplement(DNAString(mycoplasma_CP006916.3)))

mycoplasma_CP006916.3_char<-unlist(strsplit(mycoplasma_CP006916.3, ''))
reverseComplement_mycoplasma_CP006916.3<-unlist(strsplit(mycoplasma_CP006916.3, ''))

reverseComplement_mycoplasma_CP006916.3_char<-unlist(strsplit(reverseComplement_mycoplasma_CP006916.3, ''))


```



New function for physical properties calculation 

```{r dynchars function }
if(!require(ape)){stop('Code requires library "ape".\n')}
if(!require(reldna)){stop('Code requires library "reldna".\n')}
if(!require(parallel)){stop('Code requires library "parallel".\n')}
if(!require(Biostrings)){stop('Code requires library "Biostrings".\n')}

get_forward_subseq <- function(seq, tss, boundaries) {
  return(as.character(
    DNAString(seq,
              start = tss-boundaries[1],
              nchar = sum(boundaries)+1)))
}

get_reverse_subseq <- function(seq, tss, boundaries) {
  return(as.character(
    reverseComplement(
      DNAString(seq,
                start = tss-boundaries[2],
                nchar = sum(boundaries)+1))))
}

get_forward_substring <- function(seq, tss, boundaries) {
   return(seq[(tss-boundaries[1]):(tss+boundaries[2])])
}

get_reverse_substring <- function(seq, tss, boundaries) {
  res <- rev(chartr("ATGC", "TACG", seq[(tss-boundaries[2]):(tss+boundaries[1])]))
  return(res)
}

cut_char_vec <- function(char_vec, tss, boundaries, strand) {
  res <- switch(strand,
         forward = char_vec[(tss-boundaries[1]):(tss+boundaries[2])],
         reverse = char_vec[(tss-boundaries[2]):(tss+boundaries[1])]
         )
  return(res)
}

#' Title
#'
#' @param seq full sequence (chromosome)
#' @param interval requested interval
#' @param tss
#' @param strand DNA strand
#'
#' @return
#' @export
#'
#' @examples
dynchars <- function(seq, average_interval_size, tss, boundaries, strand=c('forward','reverse')) {

  if (missing(average_interval_size))
  stop("Need to specify average_interval_size")

  if(!is.character(seq))
    stop("Sequence must be a character vector containing A, C, G, T letters only")
  #make sure no small letters are present
  seq <- toupper(seq)
  # if((tss - average_interval_size < 1) || (tss + average_interval_size > length(seq)))
  #   stop("Considered average_interval_size exceeds the size of the sequence")
  
  # seq <- switch(strand, 
  #               forward = seq,
  #               reverse = as.character(reverseComplement(DNAString(seq))))
  # seq <- unlist(strsplit(seq, ''))
  # seq <- toupper(seq)
  
  boundaries <- boundaries + average_interval_size %/% 2
  
  seq <- cut_char_vec(seq, tss, boundaries, strand)

  # strand <- match.arg(strand)
  # seq <- switch(strand,
  #             forward=get_forward_substring(seq, tss, boundaries),
  #             reverse=get_forward_substring(seq, tss, boundaries))#get_reverse_substring(seq, tss, boundaries))

  a<-3.4*10^(-10)
  #I<-c(7.6, 4.8, 8.2, 4.1)*10^(-44)
  K<-c(227, 155, 220, 149)*10^(-20)
  V<-c(2.09, 1.43, 3.12, 2.12)*10^(-20)

  csA<-cumsum(seq=='A')
  csT<-cumsum(seq=='T')
  csG<-cumsum(seq=='G')
  csC<-cumsum(seq=='C')

  countA = csA[average_interval_size:length(csA)]-c(0, csA[1:(length(csA)-average_interval_size)])
  countT = csT[average_interval_size:length(csT)]-c(0, csT[1:(length(csT)-average_interval_size)])
  countG = csG[average_interval_size:length(csG)]-c(0, csG[1:(length(csG)-average_interval_size)])
  countC = csC[average_interval_size:length(csC)]-c(0, csC[1:(length(csC)-average_interval_size)])

  M <- cbind(countA, countT, countG, countC)/average_interval_size

  #Is <- as.numeric(M%*%I)#! numeric conversion
  Ks <- as.numeric(M %*% K)
  Vs <- as.numeric(M %*% V)

  E0 <- (8*(Ks*Vs)^0.5)* 6E23 / 4184
  d <- ((Ks*a^2)/Vs)^(0.5)/a;
  gc = M[,3] + M[,4]

  dynchars_return <- list(E0=E0, d=d, gc=gc)
  return(dynchars_return)
}

#' Title
#'
#' @param tss_position
#' @param extended_string -- vector of characters
#' @param ep_interval
#' @param zout
#' @param strand
#'
#' @return
#' @export
#'
#' @examples
calculate_EP_on_interval <- function(tss_position, extended_string, ep_interval=c(270, 130), zout=-540:179, strand=c('forward','reverse')) {
  #lseqspline1D(substr(e.coli_U00096.2, exp_tsss[i]-250, exp_tsss[i]+150), bound=c(50, 350), ref=251 )
  # strand<- match.arg(strand)
  #extended_string <- unlist(strsplit(extended_string, ''))
  # subseq <-switch(strand,
  #                 forward = get_forward_substring(extended_string, tss_position, ep_interval),
  #                 reverse = get_reverse_substring(extended_string, tss_position, ep_interval))
  # subseq <- paste0(subseq, collapse = "")
  
  # subseq <- switch(strand,
  #                  forward = get_forward_subseq(extended_string, tss_position, ep_interval),
  #                  reverse = get_reverse_subseq(extended_string, tss_position, ep_interval))
  # subseq <- switch(strand,
  #                  forward = get_forward_substring(extended_string, tss_position, ep_interval),
  #                  reverse = get_reverse_substring(extended_string, tss_position, ep_interval))
  # subseq <- as.character(subseq)
  subseq <- as.character(cut_char_vec(extended_string, tss_position, ep_interval, strand))
  subseq <- paste0(subseq, collapse = "")
  p <- lseqspline1D(
    subseq,
    bound=c(50, 350),
    ref=271)
  return(p$mpot[p$x %in% zout])
}

#' Title
#'
#' @param dnaSeq
#' @param dynamic_interval
#' @param tss_position
#' @param ep_interval
#' @param zout
#' @param strand
#'
#' @return
#' @export
#'
#' @examples
calculate_profile <- function(dnaSeq, average_interval_size, tss_position, dynamic_interval, ep_interval=c(270, 130), zout=-540:179, strand=c('forward','reverse')) {
  #ep_interval = c(163, 213) for reverse, for forward c(267, 217), zout - the same
  #seq, average_interval_size, tss, boundaries, strand=c('forward','reverse')
  #dnaSeq <- unlist(strsplit(dnaSeq, ''))
  dynRes <- dynchars(dnaSeq, average_interval_size, tss_position,  dynamic_interval, strand)
  epRes <- calculate_EP_on_interval(tss_position, dnaSeq, ep_interval, zout, strand)
  return(c(dynRes$E0, dynRes$d, epRes, dynRes$gc))
}

#average_interval_size 200
#dynamic_interval -150;50


#sample calculation

seq <- substr(e.coli_U00096.2, start = 1, stop = 10000)
# #seq <- 'aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaccccccccccccccccccccccccccccccccccccccccccccc'
vec_seq <- strsplit(seq, split = '')[[1]]

get_forward_subseq(seq = seq, tss = 20, boundaries = c(10, 23))

get_reverse_subseq(seq = seq, tss = 30, boundaries = c(10, 23))

as.character(cut_char_vec(vec_seq, tss = 50, boundaries = c(15,15), strand = 'reverse'))

dynchars(seq = vec_seq, average_interval_size = 200, tss = 2000, boundaries = c(150,50),strand = 'forward')


  # strand <- match.arg(strand)


calculate_EP_on_interval(tss_position = 1000, extended_string = vec_seq, ep_interval = c(350, 150), strand = 'forward') #Ep demands vector!



calculate_profile(dnaSeq = vec_seq, average_interval_size = 30, dynamic_interval = c(10, 30), tss_position = 1000, ep_interval = c(350, 150))


#reverse strand for dynchars - problems
get_forward_substring(strsplit('aaaaaaaaaaaaaaaaaaacccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccc', split = '')[[1]], tss = 20, boundaries = c(10,10))
get_reverse_substring(strsplit('aaaaaaaaaaaaaaaaaaacccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccccc', split = '')[[1]], tss = 20, boundaries = c(10,10))
```

First to set parameters for dynchars
```{r Genome-wide characteristics profiles for three bacteria, warning=FALSE, message=FALSE, cache=TRUE}
average_interval_size <- 200
dynamic_interval <- c(150, 50)

whole_e.coli_dynchars_forward <- dynchars((e.coli_U00096.2_char), average_interval_size = average_interval_size, tss = length(e.coli_U00096.2_char)/2 , boundaries = c(0,  length(e.coli_U00096.2_char)), strand = 'forward')

#for genome-wide vector using pseudoTSS at the center cause error. The vector is obtained using reversed initially genome
whole_e.coli_dynchars_reverse <- dynchars(reverseComplement_e.coli_U00096.2_char, average_interval_size = average_interval_size, tss = length(e.coli_U00096.2_char)/2 , boundaries = c(0,  length(e.coli_U00096.2_char)), strand = 'forward')

str(whole_e.coli_dynchars_forward)
str(whole_e.coli_dynchars_reverse)

whole_e.coli_EP_forward <- lseqspline1D(s = e.coli_U00096.2, bound = c(1, nchar(e.coli_U00096.2)), width = 1, ref = nchar(e.coli_U00096.2)/2)

whole_e.coli_EP_reverse <- lseqspline1D(s = reverseComplement_e.coli_U00096.2, bound = c(1, nchar(reverseComplement_e.coli_U00096.2)), width = 1, ref = nchar(reverseComplement_e.coli_U00096.2)/2)

#same for Synechocystis

whole_synechocystis_dynchars_forward <- dynchars((synechocystis_NC_000911.1_char), average_interval_size = average_interval_size, tss = length(synechocystis_NC_000911.1_char)/2 , boundaries = c(0,  length(synechocystis_NC_000911.1_char)), strand = 'forward')
whole_synechocystis_dynchars_reverse <- dynchars(reverseComplement_synechocystis_NC_000911.1_char, average_interval_size = average_interval_size, tss = length(reverseComplement_synechocystis_NC_000911.1_char)/2 , boundaries = c(0,  length(reverseComplement_synechocystis_NC_000911.1_char)), strand = 'forward')

#same for mycoplasma

whole_mycoplasma_dynchars_forward <- dynchars((mycoplasma_CP006916.3_char), average_interval_size = average_interval_size, tss = length(mycoplasma_CP006916.3_char)/2 , boundaries = c(0,  length(mycoplasma_CP006916.3_char)), strand = 'forward')
whole_mycoplasma_dynchars_reverse <- dynchars(reverseComplement_mycoplasma_CP006916.3_char, average_interval_size = average_interval_size, tss = length(reverseComplement_mycoplasma_CP006916.3_char)/2 , boundaries = c(0,  length(reverseComplement_mycoplasma_CP006916.3_char)), strand = 'forward')

plot(whole_e.coli_dynchars$E0, ty = 'l')

```


Loading data on sequences of different types (promoters, non-promoters, genes, islands, and lowscore) frm Github (or from additional .Rdata files)
Alternatively dataset_pro as data frame is available on github
```{r dataset_pro as df from Github}

# #load('/home/jane/Документы/Misha/Lab/mol_a-16/spline_dataset_pro.Rdata')
dataset_pro_df <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/dataset_pro_df.csv')
str(dataset_pro_df)
```

Extracting data on all promoters and on experimentaly found ones - including previosely calculated electrostatic potential profiles
```{r}
pro_names <- as.character(dataset_pro_df$X)

which.experimental <- sapply(dataset_pro_df$evidence, function(x) {x == 'experimental'})
exp_names <- as.character(dataset_pro_df$X[which.experimental])
exp_tsss <-dataset_pro_df$TSS[which.experimental]
exp_strands <- as.character(dataset_pro_df$strand[which.experimental])
```

# # # Changing genomic coordinate used for dataset_... - TSS is according to 5'-end for both strands 
```{r}

exp_tsss[which(exp_strands=='reverse')] <- nchar(e.coli_U00096.2)-exp_tsss[which(exp_strands=='reverse')]
#exp_tsss<--exp_tsss
```


```{r strReverse funtcion}
## a useful function: rev() for strings
strReverse <- function(x)
        sapply(lapply(strsplit(x, NULL), rev), paste, collapse="")
```


Creating matrices for data on activation energy ('E01') and size ('d1'). The matrices is 699*201 - 699 promoters sequences, 201 value for a physical property profile
to 5'-end for both strands 


Creating matrices of dynamical properties and GC-content for intervals [-150; 50] nts
```{r E. coli promoters physics, warning=FALSE, message=FALSE, cache=TRUE}
for (i in paste0('ecoli_', c('seqsforward', 'seqsreverse', 'E01forward', 'E01reverse', 'd1forward', 'd1reverse', 'GCforward', 'GCreverse', 'EPforward', 'EPreverse'))) {
  assign(i, c())
}
# #zout<- -480:239 #better use default
#halfsize of the genome to use in a fucntion run
interv <- 500
for (i in 1:length(exp_tsss)) {
  if (exp_strands[i]=='forward') {
    
    ecoli_seqsforward <- c(ecoli_seqsforward, get_forward_subseq(seq = e.coli_U00096.2, tss = exp_tsss[i], boundaries = c(150, 50)))
   # # 

    seq_tmp_forward <- strsplit(get_forward_subseq(seq = e.coli_U00096.2, tss = exp_tsss[i], boundaries = c(interv, interv-1)), split = '')[[1]] #used in profiles calculation
    
    ecoli_E01forward<-rbind(ecoli_E01forward, dynchars(seq = seq_tmp_forward , average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$E0)
    ecoli_d1forward<-rbind(ecoli_d1forward, dynchars(seq = seq_tmp_forward, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$d)
    ecoli_GCforward<-rbind(ecoli_GCforward, dynchars(seq = seq_tmp_forward, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$gc)
   
    ecoli_EPforward <- rbind(ecoli_EPforward, 
                       calculate_EP_on_interval(extended_string = seq_tmp_forward, tss_position = interv+1, strand = 'forward' ))
    
     } else {
    ecoli_seqsreverse <- c(ecoli_seqsreverse, get_reverse_subseq(seq = e.coli_U00096.2, tss = exp_tsss[i], boundaries = c(150, 50)))
    
    seq_tmp_reverse <- strsplit(get_reverse_subseq(seq = e.coli_U00096.2, tss = exp_tsss[i], boundaries = c(interv, interv-1)), split = '')[[1]] #used in profiles calculation
    
    ecoli_E01reverse<-rbind(ecoli_E01reverse, dynchars(seq = seq_tmp_reverse , average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$E0)
    ecoli_d1reverse<-rbind(ecoli_d1reverse, dynchars(seq = seq_tmp_reverse, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$d)
    ecoli_GCreverse<-rbind(ecoli_GCreverse, dynchars(seq = seq_tmp_reverse, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$gc)
   
    ecoli_EPreverse <- rbind(ecoli_EPreverse, 
                       calculate_EP_on_interval(extended_string = seq_tmp_reverse, tss_position = interv+1, strand = 'reverse' ))
    
    }
}


#merging matrices for forward and reverse strands  together
ecoli_seqs <- toupper(c(ecoli_seqsforward, ecoli_seqsreverse))
ecoli_E01<-rbind(ecoli_E01forward, ecoli_E01reverse)
#aeos2<-rbind(aeos2forward, aeos2reverse)
ecoli_d1<-rbind(ecoli_d1forward, ecoli_d1reverse)
#aeos4<-rbind(aeos4forward, aeos4reverse)
ecoli_GC <- rbind(ecoli_GCforward, ecoli_GCreverse)
ecoli_EP<-rbind(ecoli_EPforward, ecoli_EPreverse)
```




```{r Datasets from An experimentally anchored map of transcriptional start sites in the model cyanobacterium Synechocystis sp. PCC6803 Jan Mitschke, 2011}

synechocystis_proms <- (read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/synechocystis_proms_datasets_1015154108_sd01.csv', skip = 14, header = T))
colnames(synechocystis_proms) <- c('TSS', 'p.start', 'p.end', 'strand', 'read', 'observed sequence')

#to unify strands names
levels(synechocystis_proms$strand) <- c('forward', 'reverse')

```



Creating matrices of dynamical properties and GC-content for intervals [-150; 50] nts
```{r Synechocystis proms physics, warning=FALSE, message=FALSE, cache=TRUE}
for (i in paste0('synechocystis_', c('seqsforward', 'seqsreverse', 'E01forward', 'E01reverse', 'd1forward', 'd1reverse', 'GCforward', 'GCreverse', 'EPforward', 'EPreverse'))) {
  assign(i, c())
}
# #zout<- -480:239 #better use default
#halfsize of the genome to use in a fucntion run
interv <- 500
for (i in 1:length(synechocystis_proms$TSS)) {
  if (synechocystis_proms$strand[i]=='forward') {
    
    synechocystis_seqsforward <- c(synechocystis_seqsforward, get_forward_subseq(seq = synechocystis_NC_000911.1, tss = synechocystis_proms$TSS[i], boundaries = c(150, 50)))
   # # 

    seq_tmp_forward <- strsplit(get_forward_subseq(seq = synechocystis_NC_000911.1, tss = synechocystis_proms$TSS[i], boundaries = c(interv, interv-2)), split = '')[[1]] #used in profiles calculation
    
    synechocystis_E01forward<-rbind(synechocystis_E01forward, dynchars(seq = seq_tmp_forward , average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$E0)
    synechocystis_d1forward<-rbind(synechocystis_d1forward, dynchars(seq = seq_tmp_forward, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$d)
    synechocystis_GCforward<-rbind(synechocystis_GCforward, dynchars(seq = seq_tmp_forward, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$gc)
   
    synechocystis_EPforward <- rbind(synechocystis_EPforward, 
                       calculate_EP_on_interval(extended_string = seq_tmp_forward, tss_position = interv+1, strand = 'forward' ))
    
     } else {
    synechocystis_seqsreverse <- c(synechocystis_seqsreverse, get_reverse_subseq(seq = synechocystis_NC_000911.1, tss = synechocystis_proms$TSS[i], boundaries = c(150, 50)))
    
    seq_tmp_reverse <- strsplit(get_reverse_subseq(seq = synechocystis_NC_000911.1, tss = exp_tsss[i], boundaries = c(interv, interv-2)), split = '')[[1]] #used in profiles calculation
    
    synechocystis_E01reverse<-rbind(synechocystis_E01reverse, dynchars(seq = seq_tmp_reverse , average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$E0)
    synechocystis_d1reverse<-rbind(synechocystis_d1reverse, dynchars(seq = seq_tmp_reverse, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$d)
    synechocystis_GCreverse<-rbind(synechocystis_GCreverse, dynchars(seq = seq_tmp_reverse, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$gc)
   
    synechocystis_EPreverse <- rbind(synechocystis_EPreverse, 
                       calculate_EP_on_interval(extended_string = seq_tmp_reverse, tss_position = interv+1, strand = 'reverse' ))
    
    }
}


#merging matrices for forward and reverse strands  together
synechocystis_seqs <- toupper(c(synechocystis_seqsforward, synechocystis_seqsreverse))
synechocystis_E01<-rbind(synechocystis_E01forward, synechocystis_E01reverse)
#aeos2<-rbind(aeos2forward, aeos2reverse)
synechocystis_d1<-rbind(synechocystis_d1forward, synechocystis_d1reverse)
#aeos4<-rbind(aeos4forward, aeos4reverse)
synechocystis_GC <- rbind(synechocystis_GCforward, synechocystis_GCreverse)
synechocystis_EP<-rbind(synechocystis_EPforward, synechocystis_EPreverse)
```




```{r  Reading in M. gallisepticum strain S6 promoters from Fisunov et al.}

mycoplasma_proms <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/mycoplasma_gallisepticum_fisunov_2016.csv', header = T)
colnames(mycoplasma_proms)[1:2] <- c('strand', 'TSS')
#to unify strands names
tmp <- (factor(mycoplasma_proms$strand))
levels(tmp)<- c('forward', 'reverse')
mycoplasma_proms$strand <- tmp
```

Right flank needs to be extended by 250 nts
```{r Mycoplasma extension}
extended_by <- 500 #both flanks
extended_mycoplasma_CP006916.3 <- paste0(substr(mycoplasma_CP006916.3, nchar(mycoplasma_CP006916.3)-extended_by, nchar(mycoplasma_CP006916.3)),
                                         mycoplasma_CP006916.3,
                                         substr(mycoplasma_CP006916.3, 1, extended_by))
                                         

extended_mycoplasma_CP006916.3_char <- strsplit(extended_mycoplasma_CP006916.3, split = '')[[1]]
```

Creating matrices of dynamical properties and GC-content for intervals [-150; 50] nts
```{r mycoplasma proms physics, warning=FALSE, message=FALSE, cache=TRUE}
for (i in paste0('mycoplasma_', c('seqsforward', 'seqsreverse', 'E01forward', 'E01reverse', 'd1forward', 'd1reverse', 'GCforward', 'GCreverse', 'EPforward', 'EPreverse'))) {
  assign(i, c())
}
# #zout<- -480:239 #better use default
#halfsize of the genome to use in a fucntion run
interv <- 500
for (i in 1:length(mycoplasma_proms$TSS)) {
  if (mycoplasma_proms$strand[i]=='forward') {
    
    mycoplasma_seqsforward <- c(mycoplasma_seqsforward, get_forward_subseq(seq = extended_mycoplasma_CP006916.3, tss = mycoplasma_proms$TSS[i] + extended_by, boundaries = c(150, 50)))
   # #! note all TSS are shifted

    seq_tmp_forward <- strsplit(get_forward_subseq(seq = extended_mycoplasma_CP006916.3, tss = mycoplasma_proms$TSS[i]+ extended_by, boundaries = c(interv, interv-2)), split = '')[[1]] #used in profiles calculation
    
    mycoplasma_E01forward<-rbind(mycoplasma_E01forward, dynchars(seq = seq_tmp_forward , average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$E0)
    mycoplasma_d1forward<-rbind(mycoplasma_d1forward, dynchars(seq = seq_tmp_forward, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$d)
    mycoplasma_GCforward<-rbind(mycoplasma_GCforward, dynchars(seq = seq_tmp_forward, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'forward')$gc)
   
    mycoplasma_EPforward <- rbind(mycoplasma_EPforward, 
                       calculate_EP_on_interval(extended_string = seq_tmp_forward, tss_position = interv+1, strand = 'forward' ))
    
     } else {
    mycoplasma_seqsreverse <- c(mycoplasma_seqsreverse, get_reverse_subseq(seq = extended_mycoplasma_CP006916.3, tss = mycoplasma_proms$TSS[i]+ extended_by, boundaries = c(150, 50)))
    
    seq_tmp_reverse <- strsplit(get_reverse_subseq(seq = extended_mycoplasma_CP006916.3, tss = mycoplasma_proms$TSS[i]+ extended_by, boundaries = c(interv, interv-2)), split = '')[[1]] #used in profiles calculation
    
    mycoplasma_E01reverse<-rbind(mycoplasma_E01reverse, dynchars(seq = seq_tmp_reverse , average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$E0)
    mycoplasma_d1reverse<-rbind(mycoplasma_d1reverse, dynchars(seq = seq_tmp_reverse, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$d)
    mycoplasma_GCreverse<-rbind(mycoplasma_GCreverse, dynchars(seq = seq_tmp_reverse, average_interval_size = average_interval_size, tss = interv+1, boundaries = c(150, 50), strand = 'reverse')$gc)
   
    mycoplasma_EPreverse <- rbind(mycoplasma_EPreverse, 
                       calculate_EP_on_interval(extended_string = seq_tmp_reverse, tss_position = interv+1, strand = 'reverse' ))
    
    }
}


#merging matrices for forward and reverse strands  together
mycoplasma_seqs <- toupper(c(mycoplasma_seqsforward, mycoplasma_seqsreverse))
mycoplasma_E01<-rbind(mycoplasma_E01forward, mycoplasma_E01reverse)
#aeos2<-rbind(aeos2forward, aeos2reverse)
mycoplasma_d1<-rbind(mycoplasma_d1forward, mycoplasma_d1reverse)
#aeos4<-rbind(aeos4forward, aeos4reverse)
mycoplasma_GC <- rbind(mycoplasma_GCforward, mycoplasma_GCreverse)
mycoplasma_EP<-rbind(mycoplasma_EPforward, mycoplasma_EPreverse)

```


```{r Plotting characteristics profiles for 3 bacteria, eval=F}

par(mfrow = c(4,3),
    mar = rep(0.5,4))

matplot(E01, type = 'l', col = 1)
matplot(d1, type = 'l', col = 1)
matplot(GC, type = 'l', col = 1)
matplot(EP, type = 'l', col = 1)

matplot(synechocystis_E01, type = 'l', col = 1)
matplot(synechocystis_d1, type = 'l', col = 1)
matplot(synechocystis_GC, type = 'l', col = 1)
matplot(synechocystis_EP, type = 'l', col = 1)

matplot(mycoplasma_E01, type = 'l', col = 1)
matplot(mycoplasma_d1, type = 'l', col = 1)
matplot(mycoplasma_GC, type = 'l', col = 1)
matplot(mycoplasma_EP, type = 'l', col = 1)

```


```{r string distance, warning=FALSE, message=FALSE, cache=TRUE}

mycoplasma_synechocystis_e.coli_proms_strings <- sapply(c(mycoplasma_seqs, synechocystis_seqs, seqs),  function(x) {paste0(x, collapse = '')})
mycoplasma_synechocystis_e.coli_proms_DNAStringSet <- DNAStringSet(mycoplasma_synechocystis_e.coli_proms_strings)

mycoplasma_synechocystis_e.coli_proms_stringdist<- stringDist(mycoplasma_synechocystis_e.coli_proms_DNAStringSet, method = 'levenshtein')

ward_mycoplasma_synechocystis_e.coli_proms_stringdist<- hclust(mycoplasma_synechocystis_e.coli_proms_stringdist, method = 'ward.D2')

colvec_mycoplasma_synechocystis_e.coli_proms<- c(rep('magenta', length(mycoplasma_proms$Position)), rep('darkgreen', length(synechocystis_proms$X1.TSS)),rep('orange', length(exp_tsss)))

ward_mycoplasma_synechocystis_e.coli_proms_stringdist %>% as.dendrogram %>% color_branches(k = length(ward_mycoplasma_synechocystis_e.coli_proms_stringdist$order), col = colvec_mycoplasma_synechocystis_e.coli_proms[order.dendrogram(ward_mycoplasma_synechocystis_e.coli_proms_stringdist %>% as.dendrogram)]) %>% set('labels_cex', 1E-5) %>% raise.dendrogram(max(ward_mycoplasma_synechocystis_e.coli_proms_stringdist$height)*0.2) -> to_plot_ward_mycoplasma_synechocystis_e.coli_proms_stringdist


# #plot(to_plot_ward_mycoplasma_synechocystis_e.coli_proms_stringdist, main = 'E. coli, Synechocystis sp., and Mycoplasma gallisepticum promoters:\n DNA Levenshtein distance')
```



```{r three bacteria promoters mixture - all the characteristics, warning=FALSE, message=FALSE, cache=TRUE, fig.width = 10, fig.asp = 1}

cex.main = 2


par(mfrow = c(5,1),
    mar = rep(1.2,4),
    oma = rep(1.2,4))

to_plot_wards <- list()
for (i in c('E01', 'd1', 'GC', 'EP')){
  three_tmp <- rbind(get(i), get(paste0('synechocystis_', i)), get(paste0('mycoplasma_', i)))
  ward_three_tmp<- hclust.vector(t(scale(t(three_tmp), center = T, scale = T)), method = 'ward') #!!

  colvec_ward_three_tmp <- c(rep('orange', length(exp_tsss)), rep('darkgreen',length(synechocystis_proms$X1.TSS)), rep('magenta',length(mycoplasma_proms$Position)))[order.dendrogram(ward_three_tmp %>% as.dendrogram )]

  ward_three_tmp %>% as.dendrogram %>% color_branches(k = length(ward_three_tmp$order), col = colvec_ward_three_tmp) %>% set('labels_cex', 0.000001) %>% raise.dendrogram(max(ward_three_tmp$height)*.2) -> to_plot_ward_three_tmp

  plot(to_plot_ward_three_tmp, labels = F, main = paste0('E. coli, Synechocystis sp., and Mycoplasma gallisepticum promoters (', i, ')'), cex.main = cex.main )
  
  assign(paste0('to_plot_wards', i), ward_three_tmp)
}

plot(to_plot_ward_mycoplasma_synechocystis_e.coli_proms_stringdist, main = 'E. coli, Synechocystis sp., and Mycoplasma gallisepticum promoters: DNA Levenshtein distance', cex.main = cex.main )

legend('topright', legend = c('E. coli', 'Synechocystis sp.', 'Mycoplasma gallisepticum'), fill = rev(unique(colvec_mycoplasma_synechocystis_e.coli_proms)), bty = 'n')
```



```{r substring occurences, warning=FALSE, message=FALSE, cache=TRUE, fig.width = 10, fig.asp = 1}

#better way to find oligos frequnces with biostrigs
all_seqs_oligos_freq <- c()
for (i in list(seqs, synechocystis_seqs, mycoplasma_seqs, e.coli_U00096.2, synechocystis_NC_000911.1, mycoplasma_CP006916.3)){
  ###print(length(i))
  tmp <- DNAString(paste0(unlist(i), collapse = ''))
  
  #tmp_StringSet <- DNAStringSet(sapply(t(tmp), function(x) {paste0(x, collapse = '')}))

all_seqs_oligos_freq <- cbind(all_seqs_oligos_freq, oligonucleotideFrequency(tmp, 6, as.prob = T))
}



colnames(all_seqs_oligos_freq) <- c('E. coli promoters', 'Synechocystis sp. promoters', 'Mycoplasma gallisepticum promoters', 'E. coli\n complete genome', 'Synechocystis sp.\n complete genome', 'Mycoplasma gallisepticum\n complete genome')
# #View(all_seqs_oligos_freq)
#phrasetables_6 <- lapply(grep('*_ng6', x = ls(), value = T), FUN = function(x) {get.phrasetable(get(x))}) #no [1:20,]


par(mfrow = c(2,3))
ylim = range(all_seqs_oligos_freq)

for (i in 1:ncol(all_seqs_oligos_freq)){
  barplot(sort(all_seqs_oligos_freq[1:20,i], decreasing = T), las = 2, ylim = ylim, main = colnames(all_seqs_oligos_freq) [i])
}

```


```{r hexamers occurences PCA, pca 1-2, warning=FALSE, message=FALSE, cache=TRUE}
PCA <- prcomp(all_seqs_oligos_freq, scale = T)
#fviz_pca_biplot(PCA12, col.var = 2, labelsize = 0.8)+ggtitle('PCA result on hexamers occurences')

label.size <- 1.7
#library(ggfortify)
#autoplot(PCA,
    #     shape = FALSE,
   #      loadings = TRUE,
  #       label = TRUE,
 #        loadings.label = TRUE,
#         label.size = label.size)+
 # theme_bw()

fviz_pca_biplot(PCA, 
             axes = c (1,2),
             #labelsize = label.size,
             #alpha.ind="contrib",
             col.ind = 'contrib'
             ) + scale_color_gradient2(low="white", 
                                       mid="blue",
                                       high="red", 
                                       midpoint=0.8)
  
```

## hexanucleotides vs their frequences
```{r Other text representation, warning=FALSE, message=FALSE, cache=TRUE, results = 'hide'}

#panel.cor function
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
{
    usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y)) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
 
    test <- cor.test(x,y) 
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 
 
    text(0.5, 0.5, txt, cex = cex * r) 
    text(.8, .8, Signif, cex=cex, col=2) 
}

pairs(all_seqs_oligos_freq, lower.panel=panel.smooth, upper.panel=panel.cor)

```


```{r Oligonucleotide occurences, warning=FALSE, message=FALSE, cache=TRUE, fig.width = 10, fig.asp = 1}

#par(mfrow = c(2,4))
#apply(all_seqs_oligos_freq, 2, function(x) {
 # boxplot(x, ylim = range(all_seqs_oligos_freq))
#  }
 # )

for (i in 1:ncol(all_seqs_oligos_freq)){
  assign(paste0('lm', i), lm(all_seqs_oligos_freq[,i] ~ seq_along(all_seqs_oligos_freq[,i])))
}
```

```{r Descriptive statistics for hexanucleotide frequences, warning=FALSE, message=FALSE, cache=TRUE}

par(mfrow = c(6,5),
    mar = rep(2,4))
plotnames <- c('E. coli promoters', 'Synechocystis sp. promoters', 'Mycoplasma gallisepticum promoters', 'E. coli complete genome', 'Synechocystis sp. complete genome', 'Mycoplasma gallisepticum complete genome')
for (i in 1:ncol(all_seqs_oligos_freq)){
 
  plot(1, type="n", axes=F, xlab="", ylab="")
  mtext(plotnames[i], side = 3, col = 2, cex= 0.7, line = -3)

  plot(get(paste0('lm', i)))
 
}
#boxplot(all_seqs_oligos_freq[,1])
#identify(rep(1, length(all_seqs_oligos_freq[,1])), all_seqs_oligos_freq[,1], labels = #seq_along(all_seqs_oligos_freq[,1]))
```


```{r cuplcl usage}
#source('/home/mikhail/Documents/Script_2016_all/cut_and_plot_clusters.R')

#function to cut heirarchical clusterization results at given clusters numer and plot all pfrofiles from each

cuplcl <- function(data, method, k, main) {
  library(fastcluster)
  library(dendextend)
  hclust <- hclust.vector(data, method = method)
  cut_hclust<-cutree(hclust, k=k)
  m <- rbind(rep(1, k), c(2:(k+1)))
  layout(m)
  par(mar = c(1.5,1.5, 1, 1),
      oma = c(2,2,2,2))

  dend <- color_branches(as.dendrogram(hclust), col = cut_hclust[order.dendrogram(as.dendrogram(hclust))])
plot(dend%>%set('labels_cex', NA), main = main, cex.main = 1.75)
  
  ylim <- range(data)
  for (i in unique(cut_hclust[order.dendrogram(as.dendrogram(hclust))])) {
    matplot(t(data[which(cut_hclust==i),]), type='l', col=i, ylim = ylim)
 # #   print(c(i, names(which(cut_hclust==i))))
  }
}
```





#Separate characteristics for separate bacteria clusterization
```{r cluster analysis for various properties and 3 bacteria on E01, warning=FALSE, message=FALSE, cache=TRUE}

ward_ecoli_E01 <- hclust.vector(E01, method = 'ward')
ward_synechocystis_E01 <- hclust.vector(synechocystis_E01, method = 'ward')
ward_mycoplasma_E01 <- hclust.vector(mycoplasma_E01, method = 'ward')

#plot(ward_ecoli_E01, labels = F)
#plot(ward_synechocystis_E01, labels = F)
#plot(ward_mycoplasma_E01, labels = F)

cuplcl(E01, method = 'ward', k = 4, main = 'E. coli, E01')
cuplcl(synechocystis_E01, method = 'ward', k = 4, main = 'Synechocystis, E01')
cuplcl(mycoplasma_E01, method = 'ward', k = 4, main = 'Mycoplasma gallisepticum, E01')

```

```{r an interesting cluster for Mycoplasma E01, warning=FALSE, message=FALSE, cache=TRUE}

par(mfrow = c(2,1),
    mar = c(1.5, 1, 1.5,1),
    oma = rep(2,4))
`Sequence (nts)` <- -150:50
inds <- which(cutree(ward_mycoplasma_E01, k = 3) == 3)
matplot(`Sequence (nts)`, t(mycoplasma_E01[inds,]), type = 'l', col = 1, main = 'Mycoplasma gallisepticum: small (13 promoters) cluster E01 profiles', ylab = 'E01 value' )
abline(v = 0, lty = 3)

pos <- mycoplasma_proms[inds,'Position']
##mycoplasma_proms[inds,]

#the promoters genomic location
interv <- seq_along(mycoplasma_CP006916.3_char)
plot(interv, type = 'n', ylab = '', xlab = 'Sequence (nts)', yaxt = 'n', xlim = range(interv))
abline(v = pos , col = 'darkred')


#the interseting small cluster sequence

##mycoplasma_seqs[inds]

#logo for [-50;50] interval
mycoplasma_peculiar_E01cluster_short_seqs <- sapply(mycoplasma_seqs[inds], function(x) {substr(x, start = 100, stop = 200)})
ggseqlogo(mycoplasma_peculiar_E01cluster_short_seqs,  ncol = 1)+ scale_y_continuous(name="Информационнное содержание (бит)")+scale_x_discrete(name ="Последовательность относительно ТСТ (п.н.)", limits=1:100, labels = -50:50)+coord_fixed(ratio = 8)
#does the cluster representatives differ from the universe?


#logo for [-150;50] interval

ggseqlogo(mycoplasma_seqs[inds],  ncol = 1)+ scale_y_continuous(name="Информационнное содержание (бит)")+
scale_x_discrete(name ="Последовательность относительно ТСТ (п.н.)", limits=1:200, labels = -150:50)+coord_fixed(ratio = 8)
#does the cluster representatives differ from the universe?

#ape sweater plot

mycoplasma_peculiar_E01cluster_short_seqs_chars <- lapply(mycoplasma_peculiar_E01cluster_short_seqs, FUN = function(x) {strsplit(x, split = '')[[1]]})
# #image.DNAbin(as.DNAbin(mycoplasma_peculiar_E01cluster_short_seqs_chars))

#string bidtance for the cluster
range(stringDist(x = mycoplasma_peculiar_E01cluster_short_seqs, method = 'lev'))

plot(hclust(stringDist(x = mycoplasma_peculiar_E01cluster_short_seqs, method = 'lev')), labels = F)
```

```{r Same for scaled versions E01, warning=FALSE, message=FALSE, cache=TRUE}
cuplcl(t(scale(t(E01), center = T, scale = T)), method = 'ward', main = 'Scaled E01, E. coli', k = 4)
cuplcl(t(scale(t(synechocystis_E01), center = T, scale = T)), method = 'ward', main = 'Scaled E01, Synechocystis', k = 4)
cuplcl(t(scale(t(mycoplasma_E01), center = T, scale = T)), method = 'ward', main = 'Scaled E01, Mycoplasma gallisepticum', k = 4)

```

```{r cluster analysis for various properties and 3 bacteria on EP, warning=FALSE, message=FALSE, cache=TRUE} 
ward_ecoli_EP <- hclust.vector(EP, method = 'ward')
ward_synechocystis_EP <- hclust.vector(synechocystis_EP, method = 'ward')
ward_mycoplasma_EP <- hclust.vector(mycoplasma_EP, method = 'ward')

#plot(ward_ecoli_EP, labels = F)
#plot(ward_synechocystis_EP, labels = F)
#plot(ward_mycoplasma_EP, labels = F)

cuplcl(EP, method = 'ward', k = 4, main = 'E. coli, EP')
cuplcl(synechocystis_EP, method = 'ward', k = 4, main = 'Synechocystis, EP')
cuplcl(mycoplasma_EP, method = 'ward', k = 4, main = 'Mycoplasma gallisepticum, EP')
```

```{r another interesting small cluster for Mycoplasma EP, warning=FALSE, message=FALSE, cache=TRUE}

par(mfrow = c(2,1),
    mar = c(1.5, 1, 1.5,1),
    oma = rep(1.5,4))
`Sequence (angstrom)` <- -480:239
inds <- which(cutree(ward_mycoplasma_EP, k = 3) == 3)
matplot(`Sequence (angstrom)`, t(mycoplasma_EP[inds,]), type = 'l', col = 1, main = 'Mycoplasma gallisepticum: small (13 promoters) cluster EP profiles', ylab = 'EP value' )
abline(v = 0, lty = 3)

pos <- mycoplasma_proms[inds,'Position']
##mycoplasma_proms[inds,]
#the promoters genomic location
interv <- seq_along(mycoplasma_CP006916.3_char)
plot(interv, type = 'n', ylab = '', xlab = 'Sequence (nts)', yaxt = 'n', xlim = range(interv))
abline(v = pos , col = 'darkred')


#the interseting small cluster sequence

##mycoplasma_seqs[inds]

#logo for [-50;50] interval
mycoplasma_peculiar_EPcluster_short_seqs <- sapply(mycoplasma_seqs[inds], function(x) {substr(x, start = 100, stop = 200)})
ggseqlogo(mycoplasma_peculiar_EPcluster_short_seqs,  ncol = 1)+ scale_y_continuous(name="Информационнное содержание (бит)")+
scale_x_discrete(name ="Последовательность относительно ТСТ (п.н.)", limits=1:100, labels = -50:50)+coord_fixed(ratio = 8)
#does the cluster representatives differ from the universe?


#logo for [-150;50] interval

ggseqlogo(mycoplasma_seqs[inds],  ncol = 1)+ scale_y_continuous(name="Информационнное содержание (бит)")+
scale_x_discrete(name ="Последовательность относительно ТСТ (п.н.)", limits=1:200, labels = -150:50)+coord_fixed(ratio = 8)
#does the cluster representatives differ from the universe?

#string bidtance for the cluster
library(Biostrings)
range(stringDist(x = mycoplasma_peculiar_EPcluster_short_seqs, method = 'lev'))
#3 peculiar cluster promoters and their relation to heatshock expression from FIsunov et al.

mycoplasma_peculiar_EPcluster_short_seqs_chars <- lapply(mycoplasma_peculiar_EPcluster_short_seqs, FUN = function(x) {strsplit(x, split = '')[[1]]})
#small intersting clusters for Mycoplsma - on E01 and EP

par(mfrow = c(2,1))
image.DNAbin(as.DNAbin(mycoplasma_peculiar_EPcluster_short_seqs_chars), show.labels = F)
image.DNAbin(as.DNAbin(mycoplasma_peculiar_E01cluster_short_seqs_chars), show.labels = F)

```

```{r Same for scaled versions EP, warning=FALSE, message=FALSE, cache=TRUE}
cuplcl(t(scale(t(EP), center = T, scale = T)), method = 'ward', main = 'Scaled E01, E. coli', k = 4)
cuplcl(t(scale(t(synechocystis_EP), center = T, scale = T)), method = 'ward', main = 'Scaled E01, Synechocystis sp.', k = 4)
cuplcl(t(scale(t(mycoplasma_EP), center = T, scale = T)), method = 'ward', main = 'Scaled E01, Mycoplasma gallisepticum', k = 4)

```



```{r cluster analysis for various properties and 3 bacteria on GC, warning=FALSE, message=FALSE, cache=TRUE}
library(fastcluster)
ward_ecoli_GC <- hclust.vector(GC, method = 'ward')
ward_synechocystis_GC <- hclust.vector(synechocystis_GC, method = 'ward')
ward_mycoplasma_GC <- hclust.vector(mycoplasma_GC, method = 'ward')

#plot(ward_ecoli_GC, labels = F)
#plot(ward_synechocystis_GC, labels = F)
#plot(ward_mycoplasma_GC, labels = F)

cuplcl(t(GC), method = 'ward', k = 4, main = 'E. coli, GC')
cuplcl(t(synechocystis_GC), method = 'ward', k = 4, main = 'Synechocystis, GC')
cuplcl(t(mycoplasma_GC), method = 'ward', k = 4, main = 'Mycoplasma gallisepticum, GC')

```

```{r Same for scaled versions GC, warning=FALSE, message=FALSE, cache=TRUE}
cuplcl(t(scale(t(GC), center = T, scale = T)), method = 'ward', main = 'Scaled GC, E. coli', k = 4)
cuplcl(t(scale(t(synechocystis_GC), center = T, scale = T)), method = 'ward', main = 'Scaled GC, Synechocystis', k = 4)
cuplcl(t(scale(t(mycoplasma_GC), center = T, scale = T)), method = 'ward', main = 'Scaled GC, Mycoplasma gallisepticum', k = 4)

```


```{r Reading in transription units data, warning=FALSE, message=FALSE, cache=TRUE}

TUSet <- read.table('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/regulondb_TUSet_html_line_removed.txt', header = T, sep = '\t')
```

Promoter numbers and genenames
```{r E01 clusters population and GOstats, warning=FALSE, message=FALSE, cache=TRUE}

library(org.EcK12.eg.db)
cut_ward_ecoli_E01 <- cutree(ward_ecoli_E01, k = 4)

#table(cut_ward_ecoli_E01)

ecoli_E01_promoters_clusters <- c()
for (i in unique(cut_ward_ecoli_E01)) {
 assign(paste0('ecoli_E01_clust_', i),
       exp_names[which(cut_ward_ecoli_E01 ==i)])
  
  #alternatively
  tmp <- exp_names[which(cut_ward_ecoli_E01 ==i)]
  tmp1 <- cbind(rep(i, length(tmp)), tmp)
  
  ecoli_E01_promoters_clusters <- rbind(ecoli_E01_promoters_clusters, tmp1)
}


#promoters to genes (complete set)

genes_from_promoters_list <- c()
for (i in seq_along(exp_names)){
  tmp <- as.character(TUSet$Name.of.the.gene.s..contained.in.the.transcription.unit[which(TUSet$Promoter.Name == exp_names[i])])
  if(length(tmp) ==0) {tmp <- NA}
  tmp <- paste0(tmp, collapse = ',')
  genes_from_promoters_list <- c(genes_from_promoters_list, tmp)
}


#all the genes universe without repeating and 2 or more names in 1 vector member

fixed_vectors_genes_from_promoters_list <- strsplit(paste0(genes_from_promoters_list, collapse = ','), split = ',')[[1]]
fixed_vectors_genes_from_promoters_list <-fixed_vectors_genes_from_promoters_list[(fixed_vectors_genes_from_promoters_list)!='NA']

#clusters of genes as fixed vectors
for (i in unique(cut_ward_ecoli_E01)){
  genes_clust_tmp <- strsplit(paste0(genes_from_promoters_list[which(cut_ward_ecoli_E01 == i)], collapse = ','), split = ',')[[1]]
  if ('NA' %in% genes_clust_tmp) {genes_clust_tmp <- genes_clust_tmp[-which(genes_clust_tmp =='NA')]}
  assign(paste0('genes_clust', i), unique(genes_clust_tmp))
}


## Bimap interface:
# Convert the object to a list
xx <- as.list(org.EcK12.egALIAS2EG)
# Remove pathway identifiers that do not map to any entrez gene id
xx <- xx[!is.na(xx)]

#common names in universe (xx) and the samples

setdiffed <- setdiff(fixed_vectors_genes_from_promoters_list, names(xx))
fixed_vectors_genes_from_promoters_list <- fixed_vectors_genes_from_promoters_list[!fixed_vectors_genes_from_promoters_list%in%setdiffed]
entrezIds <- mget(fixed_vectors_genes_from_promoters_list, envir=org.EcK12.egALIAS2EG)
  ###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]

#common names in universe (xx) and the clusters 1-4

setdiffed_clust1 <- setdiff(genes_clust1, names(xx))
fixed_genes_clust1 <- genes_clust1[!genes_clust1%in%setdiffed_clust1]
entrezIds_clust1 <- mget(fixed_genes_clust1, envir=org.EcK12.egALIAS2EG)


setdiffed_clust2 <- setdiff(genes_clust2, names(xx))
fixed_genes_clust2 <- genes_clust2[!genes_clust2%in%setdiffed_clust2]
entrezIds_clust2 <- mget(fixed_genes_clust2, envir=org.EcK12.egALIAS2EG)


setdiffed_clust3 <- setdiff(genes_clust3, names(xx))
fixed_genes_clust3 <- genes_clust3[!genes_clust3%in%setdiffed_clust3]
entrezIds_clust3 <- mget(fixed_genes_clust3, envir=org.EcK12.egALIAS2EG)

setdiffed_clust4 <- setdiff(genes_clust4, names(xx))
fixed_genes_clust4 <- genes_clust4[!genes_clust4%in%setdiffed_clust4]
entrezIds_clust4 <- mget(fixed_genes_clust4, envir=org.EcK12.egALIAS2EG)

dfGO_ecoli_E01 <- as.data.frame(rbind(cbind(rep(1, length(entrezIds_clust1)), as.character(entrezIds_clust1)),
      cbind(rep(2, length(entrezIds_clust2)), as.character(entrezIds_clust2)),
      cbind(rep(3, length(entrezIds_clust3)), as.character(entrezIds_clust3)),
      cbind(rep(4, length(entrezIds_clust4)), as.character(entrezIds_clust4))
))

colnames(dfGO_ecoli_E01) <- c('group', 'Entrezid')
  
# #save(dfGO_ecoli_E01, file = '/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_E01.rda')
###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]
```




```{r GO statistics using GOstats for E01, warning=FALSE, message=FALSE, cache=TRUE}
library(GOstats)
library(Category)
library(DBI)
library(org.EcK12.eg.db)

# #load('/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_E01.rda')

entrezUniverse <- as.character(dfGO_ecoli_E01$Entrezid)

selectedEntrezIds_clust1 <- as.character(dfGO_ecoli_E01$Entrezid[which(dfGO_ecoli_E01$group == 1)])
selectedEntrezIds_clust2 <- as.character(dfGO_ecoli_E01$Entrezid[which(dfGO_ecoli_E01$group == 2)])
selectedEntrezIds_clust3 <- as.character(dfGO_ecoli_E01$Entrezid[which(dfGO_ecoli_E01$group == 3)])
selectedEntrezIds_clust4 <- as.character(dfGO_ecoli_E01$Entrezid[which(dfGO_ecoli_E01$group == 4)])

#setting cutoff value
hgCutoff <- 0.0005
#hgCutoff <- 0.1

#selecting onthology

ontology <- 'BP'
#ontology <- 'MF'

for (i in 1:4) {
params <- new("GOHyperGParams",
 geneIds=get(paste0('selectedEntrezIds_clust', i)),
 universeGeneIds=entrezUniverse,
 annotation="org.EcK12.eg.db",
 ontology=ontology,
 pvalueCutoff=hgCutoff,
 conditional=FALSE,
 testDirection="over")

  hgOver_clust <- hyperGTest(params)
  
  assign(x = paste0('df_summary_clust', i), summary(hgOver_clust))
}


head(df_summary_clust1,  20) #primarly nucleoside metabolisim and ion transport
head(df_summary_clust2,  20) #suflur,bicarbonic acid, aldehyde metabolism
head(df_summary_clust3,  20) #biogenesis, aerobic respiration
head(df_summary_clust4,  20) #sugar metabolism
#We would also like to perform a conditional test.  Instead of having to de ne a new
#GO-HyperGParamsinstance  by  hand,  we  can  create  a  copy  and  update  just  the  parameter  of interest.
#paramsCond <- params
#conditional(paramsCond) <- TRUE
#hgCondOver <- hyperGTest(paramsCond)


#hgOver_clust1 <- hyperGTest(params_clust1)
#hgOver_clust2 <- hyperGTest(params_clust2)
#hgOver_clust3 <- hyperGTest(params_clust3)
#hgOver_clust4 <- hyperGTest(params_clust4)


#df_clust1 <- summary(hgOver_clust1)
#df_clust2 <- summary(hgOver_clust2)
#df_clust3 <- summary(hgOver_clust3)
#df_clust4 <- summary(hgOver_clust4)

########cbind(rbind(df_clust2, df_clust3, df_clust4), (df_clust2, df_clust3, df_clust4)

#names(df_clust2)

```



```{r cluster analysis for various properties and 3 bacteria on d1, warning=FALSE, message=FALSE, cache=TRUE}
library(fastcluster)
ward_ecoli_d1 <- hclust.vector(d1, method = 'ward')
ward_synechocystis_d1 <- hclust.vector(synechocystis_d1, method = 'ward')
ward_mycoplasma_d1 <- hclust.vector(mycoplasma_d1, method = 'ward')

par(mfrow = c (1,3))

#plot(ward_ecoli_d1, labels = F)
#plot(ward_synechocystis_d1, labels = F)
#plot(ward_mycoplasma_d1, labels = F)

cuplcl(t(d1), method = 'ward', k = 4, main = 'E. coli, d1')
cuplcl(t(synechocystis_d1), method = 'ward', k = 4, main = 'Synechocystis, d1')
cuplcl(t(mycoplasma_d1), method = 'ward', k = 4, main = 'Mycoplasma gallisepticum, d1')

```

Promoter numbers and genenames
```{r d1 clusters population and GOstats, warning=FALSE, message=FALSE, cache=TRUE}
cut_ward_ecoli_d1 <- cutree(ward_ecoli_d1, k = 4)

#table(cut_ward_ecoli_d1)

ecoli_d1_promoters_clusters <- c()
for (i in unique(cut_ward_ecoli_d1)) {
 assign(paste0('ecoli_d1_clust_', i),
       exp_names[which(cut_ward_ecoli_d1 ==i)])
  
  #alternatively
  tmp <- exp_names[which(cut_ward_ecoli_d1 ==i)]
  tmp1 <- cbind(rep(i, length(tmp)), tmp)
  
  ecoli_d1_promoters_clusters <- rbind(ecoli_d1_promoters_clusters, tmp1)
}


#promoters to genes (complete set)

genes_from_promoters_list <- c()
for (i in seq_along(exp_names)){
  tmp <- as.character(TUSet$Name.of.the.gene.s..contained.in.the.transcription.unit[which(TUSet$Promoter.Name == exp_names[i])])
  if(length(tmp) ==0) {tmp <- NA}
  tmp <- paste0(tmp, collapse = ',')
  genes_from_promoters_list <- c(genes_from_promoters_list, tmp)
}


#all the genes universe without repeating and 2 jr more names in 1 vector member

fixed_vectors_genes_from_promoters_list <- strsplit(paste0(genes_from_promoters_list, collapse = ','), split = ',')[[1]]
fixed_vectors_genes_from_promoters_list <-fixed_vectors_genes_from_promoters_list[(fixed_vectors_genes_from_promoters_list)!='NA']

#clusters of genes as fixed vectors
for (i in unique(cut_ward_ecoli_d1)){
  genes_clust_tmp <- strsplit(paste0(genes_from_promoters_list[which(cut_ward_ecoli_d1 == i)], collapse = ','), split = ',')[[1]]
  if ('NA' %in% genes_clust_tmp) {genes_clust_tmp <- genes_clust_tmp[-which(genes_clust_tmp =='NA')]}
  assign(paste0('genes_clust', i), unique(genes_clust_tmp))
}


## Bimap interface:
# Convert the object to a list
xx <- as.list(org.EcK12.egALIAS2EG)
# Remove pathway identifiers that do not map to any entrez gene id
xx <- xx[!is.na(xx)]

#common names in universe (xx) and the samples

setdiffed <- setdiff(fixed_vectors_genes_from_promoters_list, names(xx))
fixed_vectors_genes_from_promoters_list <- fixed_vectors_genes_from_promoters_list[!fixed_vectors_genes_from_promoters_list%in%setdiffed]
entrezIds <- mget(fixed_vectors_genes_from_promoters_list, envir=org.EcK12.egALIAS2EG)
  ###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]

#common names in universe (xx) and the clusters 1-4

setdiffed_clust1 <- setdiff(genes_clust1, names(xx))
fixed_genes_clust1 <- genes_clust1[!genes_clust1%in%setdiffed_clust1]
entrezIds_clust1 <- mget(fixed_genes_clust1, envir=org.EcK12.egALIAS2EG)


setdiffed_clust2 <- setdiff(genes_clust2, names(xx))
fixed_genes_clust2 <- genes_clust2[!genes_clust2%in%setdiffed_clust2]
entrezIds_clust2 <- mget(fixed_genes_clust2, envir=org.EcK12.egALIAS2EG)


setdiffed_clust3 <- setdiff(genes_clust3, names(xx))
fixed_genes_clust3 <- genes_clust3[!genes_clust3%in%setdiffed_clust3]
entrezIds_clust3 <- mget(fixed_genes_clust3, envir=org.EcK12.egALIAS2EG)

setdiffed_clust4 <- setdiff(genes_clust4, names(xx))
fixed_genes_clust4 <- genes_clust4[!genes_clust4%in%setdiffed_clust4]
entrezIds_clust4 <- mget(fixed_genes_clust4, envir=org.EcK12.egALIAS2EG)

dfGO_ecoli_d1 <- as.data.frame(rbind(cbind(rep(1, length(entrezIds_clust1)), as.character(entrezIds_clust1)),
      cbind(rep(2, length(entrezIds_clust2)), as.character(entrezIds_clust2)),
      cbind(rep(3, length(entrezIds_clust3)), as.character(entrezIds_clust3)),
      cbind(rep(4, length(entrezIds_clust4)), as.character(entrezIds_clust4))
))

colnames(dfGO_ecoli_d1) <- c('group', 'Entrezid')
  
# #save(dfGO_ecoli_d1, file = '/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_d1.rda')
###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]
```




```{r GO statistics using GOstats for d1, warning=FALSE, message=FALSE, cache=TRUE}
library(GOstats)
library(Category)
library(DBI)
library(org.EcK12.eg.db)

# #load('/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_d1.rda')

entrezUniverse <- as.character(dfGO_ecoli_d1$Entrezid)

selectedEntrezIds_clust1 <- as.character(dfGO_ecoli_d1$Entrezid[which(dfGO_ecoli_d1$group == 1)])
selectedEntrezIds_clust2 <- as.character(dfGO_ecoli_d1$Entrezid[which(dfGO_ecoli_d1$group == 2)])
selectedEntrezIds_clust3 <- as.character(dfGO_ecoli_d1$Entrezid[which(dfGO_ecoli_d1$group == 3)])
selectedEntrezIds_clust4 <- as.character(dfGO_ecoli_d1$Entrezid[which(dfGO_ecoli_d1$group == 4)])

#setting cutoff value
hgCutoff <- 0.0005
#hgCutoff <- 0.1

#selecting onthology

ontology <- 'BP'
#ontology <- 'MF'

for (i in 1:4) {
params <- new("GOHyperGParams",
 geneIds=get(paste0('selectedEntrezIds_clust', i)),
 universeGeneIds=entrezUniverse,
 annotation="org.EcK12.eg.db",
 ontology=ontology,
 pvalueCutoff=hgCutoff,
 conditional=FALSE,
 testDirection="over")

  hgOver_clust <- hyperGTest(params)
  
  assign(x = paste0('df_summary_clust', i), summary(hgOver_clust))
}


#View(df_summary_clust1) #-
#View(df_summary_clust2) #ion transport
#View(df_summary_clust3) # citrate cycle, monosacharide
# #View(df_summary_clust4) #lipid metabolism


head(df_summary_clust1,  20)  #-
head(df_summary_clust2,  20)#ion transport
head(df_summary_clust3,  20)# citrate cycle, monosacharide
head(df_summary_clust4,  20) #lipid metabolism

#We would also like to perform a conditional test.  Instead of having to de ne a new
#GO-HyperGParamsinstance  by  hand,  we  can  create  a  copy  and  update  just  the  parameter  of interest.
#paramsCond <- params
#conditional(paramsCond) <- TRUE
#hgCondOver <- hyperGTest(paramsCond)


#hgOver_clust1 <- hyperGTest(params_clust1)
#hgOver_clust2 <- hyperGTest(params_clust2)
#hgOver_clust3 <- hyperGTest(params_clust3)
#hgOver_clust4 <- hyperGTest(params_clust4)


#df_clust1 <- summary(hgOver_clust1)
#df_clust2 <- summary(hgOver_clust2)
#df_clust3 <- summary(hgOver_clust3)
#df_clust4 <- summary(hgOver_clust4)

########cbind(rbind(df_clust2, df_clust3, df_clust4), (df_clust2, df_clust3, df_clust4)

#names(df_clust2)

```


Promoter numbers and genenames
```{r GC, clusters population and GOstats, warning=FALSE, message=FALSE, cache=TRUE}
cut_ward_ecoli_GC <- cutree(ward_ecoli_GC, k = 4)

#table(cut_ward_ecoli_GC)

ecoli_GC_promoters_clusters <- c()
for (i in unique(cut_ward_ecoli_GC)) {
 assign(paste0('ecoli_GC_clust_', i),
       exp_names[which(cut_ward_ecoli_GC ==i)])
  
  #alternatively
  tmp <- exp_names[which(cut_ward_ecoli_GC ==i)]
  tmp1 <- cbind(rep(i, length(tmp)), tmp)
  
  ecoli_GC_promoters_clusters <- rbind(ecoli_GC_promoters_clusters, tmp1)
}


#promoters to genes (complete set)

genes_from_promoters_list <- c()
for (i in seq_along(exp_names)){
  tmp <- as.character(TUSet$Name.of.the.gene.s..contained.in.the.transcription.unit[which(TUSet$Promoter.Name == exp_names[i])])
  if(length(tmp) ==0) {tmp <- NA}
  tmp <- paste0(tmp, collapse = ',')
  genes_from_promoters_list <- c(genes_from_promoters_list, tmp)
}


#all the genes universe without repeating and 2 jr more names in 1 vector member

fixed_vectors_genes_from_promoters_list <- strsplit(paste0(genes_from_promoters_list, collapse = ','), split = ',')[[1]]
fixed_vectors_genes_from_promoters_list <-fixed_vectors_genes_from_promoters_list[(fixed_vectors_genes_from_promoters_list)!='NA']

#clusters of genes as fixed vectors
for (i in unique(cut_ward_ecoli_GC)){
  genes_clust_tmp <- strsplit(paste0(genes_from_promoters_list[which(cut_ward_ecoli_GC == i)], collapse = ','), split = ',')[[1]]
  if ('NA' %in% genes_clust_tmp) {genes_clust_tmp <- genes_clust_tmp[-which(genes_clust_tmp =='NA')]}
  assign(paste0('genes_clust', i), unique(genes_clust_tmp))
}


## Bimap interface:
# Convert the object to a list
xx <- as.list(org.EcK12.egALIAS2EG)
# Remove pathway identifiers that do not map to any entrez gene id
xx <- xx[!is.na(xx)]

#common names in universe (xx) and the samples

setdiffed <- setdiff(fixed_vectors_genes_from_promoters_list, names(xx))
fixed_vectors_genes_from_promoters_list <- fixed_vectors_genes_from_promoters_list[!fixed_vectors_genes_from_promoters_list%in%setdiffed]
entrezIds <- mget(fixed_vectors_genes_from_promoters_list, envir=org.EcK12.egALIAS2EG)
  ###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]

#common names in universe (xx) and the clusters 1-4

setdiffed_clust1 <- setdiff(genes_clust1, names(xx))
fixed_genes_clust1 <- genes_clust1[!genes_clust1%in%setdiffed_clust1]
entrezIds_clust1 <- mget(fixed_genes_clust1, envir=org.EcK12.egALIAS2EG)


setdiffed_clust2 <- setdiff(genes_clust2, names(xx))
fixed_genes_clust2 <- genes_clust2[!genes_clust2%in%setdiffed_clust2]
entrezIds_clust2 <- mget(fixed_genes_clust2, envir=org.EcK12.egALIAS2EG)


setdiffed_clust3 <- setdiff(genes_clust3, names(xx))
fixed_genes_clust3 <- genes_clust3[!genes_clust3%in%setdiffed_clust3]
entrezIds_clust3 <- mget(fixed_genes_clust3, envir=org.EcK12.egALIAS2EG)

setdiffed_clust4 <- setdiff(genes_clust4, names(xx))
fixed_genes_clust4 <- genes_clust4[!genes_clust4%in%setdiffed_clust4]
entrezIds_clust4 <- mget(fixed_genes_clust4, envir=org.EcK12.egALIAS2EG)

dfGO_ecoli_GC <- as.data.frame(rbind(cbind(rep(1, length(entrezIds_clust1)), as.character(entrezIds_clust1)),
      cbind(rep(2, length(entrezIds_clust2)), as.character(entrezIds_clust2)),
      cbind(rep(3, length(entrezIds_clust3)), as.character(entrezIds_clust3)),
      cbind(rep(4, length(entrezIds_clust4)), as.character(entrezIds_clust4))
))

colnames(dfGO_ecoli_GC) <- c('group', 'Entrezid')
  
# #save(dfGO_ecoli_GC, file = '/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_GC.rda')
###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]
```




```{r GO statistics using GOstats for GC, warning=FALSE, message=FALSE, cache=TRUE}
library(GOstats)
library(Category)
library(DBI)
library(org.EcK12.eg.db)

# #load('/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_GC.rda')

entrezUniverse <- as.character(dfGO_ecoli_GC$Entrezid)

selectedEntrezIds_clust1 <- as.character(dfGO_ecoli_GC$Entrezid[which(dfGO_ecoli_GC$group == 1)])
selectedEntrezIds_clust2 <- as.character(dfGO_ecoli_GC$Entrezid[which(dfGO_ecoli_GC$group == 2)])
selectedEntrezIds_clust3 <- as.character(dfGO_ecoli_GC$Entrezid[which(dfGO_ecoli_GC$group == 3)])
selectedEntrezIds_clust4 <- as.character(dfGO_ecoli_GC$Entrezid[which(dfGO_ecoli_GC$group == 4)])

#setting cutoff value
hgCutoff <- 0.0005
#hgCutoff <- 0.1

#selecting onthology

ontology <- 'BP'
#ontology <- 'MF'

for (i in 1:4) {
params <- new("GOHyperGParams",
 geneIds=get(paste0('selectedEntrezIds_clust', i)),
 universeGeneIds=entrezUniverse,
 annotation="org.EcK12.eg.db",
 ontology=ontology,
 pvalueCutoff=hgCutoff,
 conditional=FALSE,
 testDirection="over")

  hgOver_clust <- hyperGTest(params)
  
  assign(x = paste0('df_summary_clust', i), summary(hgOver_clust))
}


#View(df_summary_clust1) #-
#View(df_summary_clust2) #ion transport
#View(df_summary_clust3) # citrate cycle, monosacharide
# #View(df_summary_clust4) #lipid metabolism


head(df_summary_clust1,  20) # alcohol metabolic process, translation
head(df_summary_clust2,  20) #organic acid and sulfur
head(df_summary_clust3,  20)  # tricarboxylic acid cycle
head(df_summary_clust4,  20)  # carbohydrates metabolism

#We would also like to perform a conditional test.  Instead of having to de ne a new
#GO-HyperGParamsinstance  by  hand,  we  can  create  a  copy  and  update  just  the  parameter  of interest.
#paramsCond <- params
#conditional(paramsCond) <- TRUE
#hgCondOver <- hyperGTest(paramsCond)


#hgOver_clust1 <- hyperGTest(params_clust1)
#hgOver_clust2 <- hyperGTest(params_clust2)
#hgOver_clust3 <- hyperGTest(params_clust3)
#hgOver_clust4 <- hyperGTest(params_clust4)


#df_clust1 <- summary(hgOver_clust1)
#df_clust2 <- summary(hgOver_clust2)
#df_clust3 <- summary(hgOver_clust3)
#df_clust4 <- summary(hgOver_clust4)

########cbind(rbind(df_clust2, df_clust3, df_clust4), (df_clust2, df_clust3, df_clust4)

#names(df_clust2)

```


Promoter numbers and genenames
```{r EP, clusters population and GOstats, warning=FALSE, message=FALSE, cache=TRUE}
cut_ward_ecoli_EP <- cutree(ward_ecoli_EP, k = 4)

#table(cut_ward_ecoli_EP)

ecoli_EP_promoters_clusters <- c()
for (i in unique(cut_ward_ecoli_EP)) {
 assign(paste0('ecoli_EP_clust_', i),
       exp_names[which(cut_ward_ecoli_EP ==i)])
  
  #alternatively
  tmp <- exp_names[which(cut_ward_ecoli_EP ==i)]
  tmp1 <- cbind(rep(i, length(tmp)), tmp)
  
  ecoli_EP_promoters_clusters <- rbind(ecoli_EP_promoters_clusters, tmp1)
}


#promoters to genes (complete set)

genes_from_promoters_list <- c()
for (i in seq_along(exp_names)){
  tmp <- as.character(TUSet$Name.of.the.gene.s..contained.in.the.transcription.unit[which(TUSet$Promoter.Name == exp_names[i])])
  if(length(tmp) ==0) {tmp <- NA}
  tmp <- paste0(tmp, collapse = ',')
  genes_from_promoters_list <- c(genes_from_promoters_list, tmp)
}


#all the genes universe without repeating and 2 jr more names in 1 vector member

fixed_vectors_genes_from_promoters_list <- strsplit(paste0(genes_from_promoters_list, collapse = ','), split = ',')[[1]]
fixed_vectors_genes_from_promoters_list <-fixed_vectors_genes_from_promoters_list[(fixed_vectors_genes_from_promoters_list)!='NA']

#clusters of genes as fixed vectors
for (i in unique(cut_ward_ecoli_EP)){
  genes_clust_tmp <- strsplit(paste0(genes_from_promoters_list[which(cut_ward_ecoli_EP == i)], collapse = ','), split = ',')[[1]]
  if ('NA' %in% genes_clust_tmp) {genes_clust_tmp <- genes_clust_tmp[-which(genes_clust_tmp =='NA')]}
  assign(paste0('genes_clust', i), unique(genes_clust_tmp))
}


## Bimap interface:
# Convert the object to a list
xx <- as.list(org.EcK12.egALIAS2EG)
# Remove pathway identifiers that do not map to any entrez gene id
xx <- xx[!is.na(xx)]

#common names in universe (xx) and the samples

setdiffed <- setdiff(fixed_vectors_genes_from_promoters_list, names(xx))
fixed_vectors_genes_from_promoters_list <- fixed_vectors_genes_from_promoters_list[!fixed_vectors_genes_from_promoters_list%in%setdiffed]
entrezIds <- mget(fixed_vectors_genes_from_promoters_list, envir=org.EcK12.egALIAS2EG)
  ###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]

#common names in universe (xx) and the clusters 1-4

setdiffed_clust1 <- setdiff(genes_clust1, names(xx))
fixed_genes_clust1 <- genes_clust1[!genes_clust1%in%setdiffed_clust1]
entrezIds_clust1 <- mget(fixed_genes_clust1, envir=org.EcK12.egALIAS2EG)


setdiffed_clust2 <- setdiff(genes_clust2, names(xx))
fixed_genes_clust2 <- genes_clust2[!genes_clust2%in%setdiffed_clust2]
entrezIds_clust2 <- mget(fixed_genes_clust2, envir=org.EcK12.egALIAS2EG)


setdiffed_clust3 <- setdiff(genes_clust3, names(xx))
fixed_genes_clust3 <- genes_clust3[!genes_clust3%in%setdiffed_clust3]
entrezIds_clust3 <- mget(fixed_genes_clust3, envir=org.EcK12.egALIAS2EG)

setdiffed_clust4 <- setdiff(genes_clust4, names(xx))
fixed_genes_clust4 <- genes_clust4[!genes_clust4%in%setdiffed_clust4]
entrezIds_clust4 <- mget(fixed_genes_clust4, envir=org.EcK12.egALIAS2EG)

dfGO_ecoli_EP <- as.data.frame(rbind(cbind(rep(1, length(entrezIds_clust1)), as.character(entrezIds_clust1)),
      cbind(rep(2, length(entrezIds_clust2)), as.character(entrezIds_clust2)),
      cbind(rep(3, length(entrezIds_clust3)), as.character(entrezIds_clust3)),
      cbind(rep(4, length(entrezIds_clust4)), as.character(entrezIds_clust4))
))

colnames(dfGO_ecoli_EP) <- c('group', 'Entrezid')
  
# #save(dfGO_ecoli_EP, file = '/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_EP.rda')
###   haveEntrezId <- names(entrezIds)[sapply(entrezIds, function(x) !is.na(x))]
```




```{r GO statistics using GOstats for EP, warning=FALSE, message=FALSE, cache=TRUE}
library(GOstats)
library(Category)
library(DBI)
library(org.EcK12.eg.db)

# #load('/home/jane/Документы/Misha/mol_a_2018/dfGO_ecoli_EP.rda')

entrezUniverse <- as.character(dfGO_ecoli_EP$Entrezid)

selectedEntrezIds_clust1 <- as.character(dfGO_ecoli_EP$Entrezid[which(dfGO_ecoli_EP$group == 1)])
selectedEntrezIds_clust2 <- as.character(dfGO_ecoli_EP$Entrezid[which(dfGO_ecoli_EP$group == 2)])
selectedEntrezIds_clust3 <- as.character(dfGO_ecoli_EP$Entrezid[which(dfGO_ecoli_EP$group == 3)])
selectedEntrezIds_clust4 <- as.character(dfGO_ecoli_EP$Entrezid[which(dfGO_ecoli_EP$group == 4)])

#setting cutoff value
hgCutoff <- 0.0005
#hgCutoff <- 0.1

#selecting onthology

ontology <- 'BP'
#ontology <- 'MF'

for (i in 1:4) {
params <- new("GOHyperGParams",
 geneIds=get(paste0('selectedEntrezIds_clust', i)),
 universeGeneIds=entrezUniverse,
 annotation="org.EcK12.eg.db",
 ontology=ontology,
 pvalueCutoff=hgCutoff,
 conditional=FALSE,
 testDirection="over")

  hgOver_clust <- hyperGTest(params)
  
  assign(x = paste0('df_summary_clust', i), summary(hgOver_clust))
}


#View(df_summary_clust1) #-
#View(df_summary_clust2) #ion transport
#View(df_summary_clust3) # citrate cycle, monosacharide
# #View(df_summary_clust4) #lipid metabolism


head(df_summary_clust1,  20)  #-
head(df_summary_clust2,  20)#ion transport
head(df_summary_clust3,  20)# citrate cycle, monosacharide
head(df_summary_clust4,  20) #lipid metabolism

#We would also like to perform a conditional test.  Instead of having to de ne a new
#GO-HyperGParamsinstance  by  hand,  we  can  create  a  copy  and  update  just  the  parameter  of interest.
#paramsCond <- params
#conditional(paramsCond) <- TRUE
#hgCondOver <- hyperGTest(paramsCond)


#hgOver_clust1 <- hyperGTest(params_clust1)
#hgOver_clust2 <- hyperGTest(params_clust2)
#hgOver_clust3 <- hyperGTest(params_clust3)
#hgOver_clust4 <- hyperGTest(params_clust4)


#df_clust1 <- summary(hgOver_clust1)
#df_clust2 <- summary(hgOver_clust2)
#df_clust3 <- summary(hgOver_clust3)
#df_clust4 <- summary(hgOver_clust4)

########cbind(rbind(df_clust2, df_clust3, df_clust4), (df_clust2, df_clust3, df_clust4)

#names(df_clust2)

```
```{r Data for GO, eval=FALSE }

library(org.EcK12.eg.db)

#not found in regulon TUset
 # absent_genenames <- c('cpxQ', 'insD', 'efeU_1', "efeU_2", "sroD", 'tcyJ', "gadF", "istR-1","istR-2", "phnE_1", "phnE_2", "dgcT", "ilvG_1", "ilvG_2"  )

  
Entrezid <- c()
group <- c()

LIST <- list(genes_clust1, genes_clust2, genes_clust3, genes_clust4)
names(LIST) <- c('cl1', 'cl2', 'cl3', 'cl4')
for ( i in  seq_along(LIST)) {
#  if ('TRUE' %in% unique(absent_genenames %in% LIST[[i]])) {
  #  tmp <- LIST[[i]][-which(LIST[[i]] %in% absent_genenames)]
 # }
  
  #handling absent genes
  for (gene in LIST[[i]]){
     get(x = gene, envir=org.EcK12.eg.db::org.EcK12.egALIAS2EG)
  }
  mget(x= as.character(LIST[[i]]), envir=org.EcK12.eg.db::org.EcK12.egALIAS2EG)
  Entrezid <- c(Entrezid, unlist(mget(x= as.character(tmp), envir=org.EcK12.eg.db::org.EcK12.egALIAS2EG)))
  group <- c(group,  rep(i, length(unlist(mget(x= as.character(tmp), envir=org.EcK12.eg.db::org.EcK12.egALIAS2EG)))))
 # dfGO <- rbind(dfGO, cbind(Entrezid, group))
}

dfGO <- data.frame(Entrezid = Entrezid, group = group)
                   
                   
#colnames(dfGO) <- c('Entrezids', 'group')
save(dfGO, file = '/home/jane/Документы/Misha/mol_a_2018/different_bacteria_physics_an_text/ecoli_sidd_clustering_df_to_go.rda')


genes_formula <- compareCluster(Entrezid~group, data=dfGO, fun='groupGO', OrgDb='org.EcK12.eg.db')
genes_formula@compareClusterResult

```
##SIDD
loading previously saved SIDD profiles
```{r SIDD clusterization, warning=FALSE, message=FALSE, cache=TRUE}

sidds_ecoli <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/sidds_ecoli.csv')[,-1] #rownumbers in column1

sidds_syn <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/sidds_synechocystis.csv')[,-1] #rownumbers in column1



library(fastcluster)
synechocystis_e.coli_proms_sidds <- cbind(sidds_ecoli, sidds_syn)


ward_synechocystis_e.coli_proms_sidd <- hclust.vector(t(synechocystis_e.coli_proms_sidds), method = 'ward')

#plot(ward_synechocystis_e.coli_proms_sidd, labels = F)

colvec_ward_synechocystis_e.coli_proms_sidd <- c(rep('darkgreen', length(synechocystis_proms$X1.TSS)),
rep('orange', length(exp_tsss)))
ward_synechocystis_e.coli_proms_sidd %>% as.dendrogram %>% color_branches(k = length(ward_synechocystis_e.coli_proms_sidd$order), col = colvec_ward_synechocystis_e.coli_proms_sidd[order.dendrogram(ward_synechocystis_e.coli_proms_sidd %>% as.dendrogram)]) %>% set('labels_cex', 1E-5) %>% raise.dendrogram(max(ward_synechocystis_e.coli_proms_sidd$height)*0.2) -> to_plot_ward_synechocystis_e.coli_proms_sidd #circlize_dendrogram(labels=F, dend_track_height = 0.95) 
  

plot(to_plot_ward_synechocystis_e.coli_proms_sidd, labels = F, main = 'E. coli and Synechocystis sp. promoters, SIDD profiles')
legend('topright', legend = c('E. coli', 'Synechocystis sp.'), fill = c('darkgreen', 'orange'))

```


```{r #tanglegrams3, cache=T}

#labels(to_plot_wards[[1]]) <- labels(to_plot_wards[[2]]) <- as.character(labels(to_plot_wards[[2]]))


#for (i in c('E01', 'd1', 'GC', 'EP')){
#for (j in c('E01', 'd1', 'GC', 'EP')){

 #   if (i == j) {next} else {
      
#tanglegram(to_plot_wardsE01, to_plot_wardsd1, lwd = 0.2, lab.cex = 1E-10)
#tanglegram(to_plot_wardsE01, to_plot_wardsGC, lwd = 0.2, lab.cex = 1E-10)
#tanglegram(to_plot_wardsE01, to_plot_wardsEP, lwd = 0.2, lab.cex = 1E-10)


#tanglegram(to_plot_wardsd1, to_plot_wardsGC, lwd = 0.2, lab.cex = 1E-10)
#tanglegram(to_plot_wardsd1, to_plot_wardsEP, lwd = 0.2, lab.cex = 1E-10)

#tanglegram(to_plot_wardsGC, to_plot_wardsEP, lwd = 0.2, lab.cex = 1E-10)
  #plot(to_plot_ward_both_tmp, labels = F, main = paste0('E. coli and Synechocystis promoters (', i, ')'), cex.main = cex.main )
  #  }
  #}
#}
```


```{r E. coli 1500 nts long sidd, warning=FALSE, message=FALSE, cache=TRUE}
# sidds_syn <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/sidds_synechocystis_1500nts.csv' )[,-1]

matplot(sidds_ecoli,ty='l',lty =1, col=1, lwd =0.5, main = 'SIDD profiles, E. coli')

ward_ecoli_1500sidd<- hclust.vector(t(sidds_ecoli), method = 'ward')
#plot(ward_ecoli_1500sidd, labels = F)
cut_ecoli_1500sidd <- cutree(ward_ecoli_1500sidd, k = 4)

cuplcl(data = t(sidds_ecoli), method = 'ward', main = 'SIDD E. coli', k = 4)
#`Sequence (nts)` <- -750:750
#par(mfrow = c(3,2))
#plot(ward_synechocystis_1500sidd, labels = F)
#for (i in unique(cut_synechocystis_1500sidd)){
  #matplot(`Sequence (nts)`, sidds_syn[, which(cut_synechocystis_1500sidd==i)], type = 'l', lty = 1, col=i)
 # abline(v = -10:10*100, lty =3, col = 'darkgrey', lwd =2)
#}
```


```{r Synechocystis 1500 nts long sidd, warning=FALSE, message=FALSE, cache=TRUE}
# sidds_syn <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/sidds_synechocystis_1500nts.csv' )[,-1]

matplot(sidds_syn,ty='l',lty =1, col=1, lwd =0.5, main = 'SIDD, Synechocystis sp.')

ward_synechocystis_1500sidd<- hclust.vector(t(sidds_syn), method = 'ward')
#plot(ward_synechocystis_1500sidd, labels = F)
cut_synechocystis_1500sidd <- cutree(ward_synechocystis_1500sidd, k = 4)

cuplcl(data = t(sidds_syn), method = 'ward', main = 'SIDD Synechocystis sp.', k = 4)
#`Sequence (nts)` <- -750:750
#par(mfrow = c(3,2))
#plot(ward_synechocystis_1500sidd, labels = F)
#for (i in unique(cut_synechocystis_1500sidd)){
  #matplot(`Sequence (nts)`, sidds_syn[, which(cut_synechocystis_1500sidd==i)], type = 'l', lty = 1, col=i)
 # abline(v = -10:10*100, lty =3, col = 'darkgrey', lwd =2)
#}
```



```{r Whole-genome SIDD vectors, warning=FALSE, message=FALSE, cache=TRUE, eval = FALSE}
load('/home/jane/Документы/Misha/mol_a_2018/whole_e.coli_sidd_2500_500.rda')
load('/home/jane/Документы/Misha/mol_a_2018/whole_e.coli_sidd_5000_1000.rda')
load('/home/jane/Документы/Misha/Lab/mol_a-16/spline_dataset_pro.Rdata')

exp_pro_numbers<- which(sapply(dataset_pro, function(x) {x$evidence =='experimental' })==T)
exp_pro_tsss <- sapply(dataset_pro, function(x){x$tss})[exp_pro_numbers]


par(mfrow = c(2,1))
interv <- 1:100000
plot(whole_e.coli_sidd_2500_500[interv], type = 'l')
abline(v = exp_pro_tsss, col = 2 )
plot(whole_e.coli_sidd_5000_1000[interv], type = 'l')
abline(v = exp_pro_tsss, col = 2 )

exp_pro_sidd_2500_500 <- sapply(exp_pro_tsss, function(x) {whole_e.coli_sidd_2500_500[(x-1000):(x+1000)]})
matplot(exp_pro_sidd_2500_500, type = 'l', col =1, lwd =0.3)

exp_pro_sidd_5000_1000 <- sapply(exp_pro_tsss, function(x) {whole_e.coli_sidd_5000_1000[(x-1000):(x+1000)]})
matplot(exp_pro_sidd_5000_1000, type = 'l', col =1, lwd =0.3)
```


```{r Rollaply max, warning=FALSE, message=FALSE, cache=TRUE, eval = FALSE}
library(zoo)

rollmax_whole_e.coli_sidd_2500_500 <- rollmax(x = whole_e.coli_sidd_2500_500, k = 250)

rollmax_whole_e.coli_sidd_5000_1000 <- rollmax(x = whole_e.coli_sidd_5000_1000, k = 250)


par(mfrow = c(2,1),
    mar = rep(0.7,4))
interv <- 1:50000
plot(rollmax_whole_e.coli_sidd_2500_500[interv], type = 'l')
abline(v = exp_pro_tsss, col = 2 )
plot(rollmax_whole_e.coli_sidd_5000_1000[interv], type = 'l')
abline(v = exp_pro_tsss, col = 2 )

which_rollmax_whole_e.coli_sidd_2500_500_0.95 <- which(rollmax_whole_e.coli_sidd_2500_500>0.95)
which_rollmax_whole_e.coli_sidd_5000_1000_0.95 <- which(rollmax_whole_e.coli_sidd_5000_1000>0.95)                                              
#how big is a moltem part of the genome (nts)
length(which_rollmax_whole_e.coli_sidd_2500_500_0.95)
length(which_rollmax_whole_e.coli_sidd_5000_1000_0.95)

par(mfrow = c(2,1),
    mar = rep(0.7,4))
interv <- 1:500000
plot(rollmax_whole_e.coli_sidd_2500_500[interv], type = 'n')
points(x = which_rollmax_whole_e.coli_sidd_2500_500_0.95, y = rep(0.95, length(which_rollmax_whole_e.coli_sidd_2500_500_0.95)), pch = '|', col = 2)
abline(v = exp_pro_tsss, col = 'darkgreen' )

plot(rollmax_whole_e.coli_sidd_5000_1000[interv], type = 'n')
points(x = which_rollmax_whole_e.coli_sidd_5000_1000_0.95, y = rep(0.95, length(which_rollmax_whole_e.coli_sidd_5000_1000_0.95)), pch = '|', col = 2)
abline(v = exp_pro_tsss, col = 'darkgreen' )

#which regions are melted for different split sets
interv <- 300000:400000
plot(rollmax_whole_e.coli_sidd_2500_500[interv], type = 'n')
points(x = which_rollmax_whole_e.coli_sidd_2500_500_0.95, y = rep(0.95, length(which_rollmax_whole_e.coli_sidd_2500_500_0.95)), pch = '|', col = 2)
text(x = 0, y =0.95, labels = '2500-500\n fragments', cex = 0.7)
points(x = which_rollmax_whole_e.coli_sidd_5000_1000_0.95, y = rep(0.75, length(which_rollmax_whole_e.coli_sidd_5000_1000_0.95)), pch = '|', col = 2)
text(x = 0, y =0.75, labels = '5000-1000\n fragments', cex = 0.7)
points(x = exp_pro_tsss, y = rep(0.55, length(exp_pro_tsss)), pch = '|', col = 'darkblue')
text(x = 0, y =0.55, labels = 'Experimentally\n evidenced\n promoters', cex = 0.7)

```
Column 1 contains numbers and is removed 
```{r CHecking bimodality for SIDD for all promoters, warning=FALSE, message=FALSE, cache=TRUE, eval = FALSE}

interv <- 750:2250
interv <- 500:2500
####interv <- 900:2100
####interv <- 1000:2000

sums_sidds_ecoli <- apply(sidds_ecoli[,-1], 2, function(x) {sum(x[interv])})

sums_sidds_syn <- apply(sidds_syn[,-1], 2, function(x) {sum(x[interv])})


#svg('/home/jane/Документы/Misha/mol_a_2018/different_bacteria_physics_an_text/densities_auc_sidd_2_bacteria.svg', height = 10, width = 10)
par(mfrow = c(2,1))
hist(sums_sidds_ecoli, breaks = 15, freq = F, xlab = 'AUC for 1500 nts-long SIDD profile', main = 'Histogram of E. coli Promoters AUC SIDD Data\n with Kernel Density Plot')
lines(density(sums_sidds_ecoli, na.rm = T))


hist(sums_sidds_syn, breaks = 15, freq = F, xlab = 'AUC for 1500 nts-long SIDD profile', main = 'Histogram of Synechocystis sp. Promoters AUC SIDD Data\n with Kernel Density Plot')
lines(density(sums_sidds_syn, na.rm = T))

#dev.off()

library(diptest)
diptest_sums_sidds_ecoli <- dip.test(sums_sidds_ecoli, simulate.p.value = T, B = 20000) #D = 0.055838, p-value < 2.2e-16 for 1500 nts -long interval
# D = 0.029814, p-value = 1e-04 for 2000 nts
diptest_sums_sidds_ecoli

diptest_sums_sidds_syn<- dip.test(sums_sidds_syn, simulate.p.value = T, B = 20000) #D = 0.037924, p-value < 2.2e-16 for 1500 nts -long interval
#D = 0.045919, p-value < 2.2e-16 for 2000 nts
diptest_sums_sidds_syn
```


```{r if dynamic properties ditributed unimodally, eval=FALSE}


funcs <- c('sum', 'mean', 'median', 'max')

for (i in funcs){
  assign(paste0(i,' _E01_e.coli_promoters'), apply(E01, MARGIN = 1, FUN = i))
}


par(mfrow = c(2,2))
for (i in funcs){
  plot(density(get(paste0(i,'_E01_e.coli_promoters'))))
}


for (i in funcs){
  assign(paste0(i,'_d1_e.coli_promoters'), apply(d1, MARGIN = 1, FUN = i))
}


par(mfrow = c(2,2))
for (i in funcs){
  plot(density(get(paste0(i,'_d1_e.coli_promoters'))))
}


for (i in funcs){
  assign(paste0(i,'_GC_e.coli_promoters'), apply(GC, MARGIN = 1, FUN = i))
}


par(mfrow = c(2,2))
for (i in funcs){
  plot(density(get(paste0(i,'_GC_e.coli_promoters'))))
}

for (i in funcs){
  assign(paste0(i,'_EP_e.coli_promoters'), apply(GC, MARGIN = 1, FUN = i))
}


par(mfrow = c(2,2))
for (i in funcs){
  plot(density(get(paste0(i,'_GC_e.coli_promoters'))))
}


```

```{r GO statistics using GOstats, warning=FALSE, message=FALSE, cache=TRUE}
library(GOstats)
library(Category)
library(DBI)
library(org.EcK12.eg.db)

# #load('/home/jane/Документы/Misha/mol_a_2018/different_bacteria_physics_an_text/ecoli_sidd_clustering_df_to_go.rda')

entrezUniverse <- as.character(dfGO$Entrezid)
selectedEntrezIds_clust2 <- as.character(dfGO$Entrezid[which(dfGO$group == 2)])
selectedEntrezIds_clust3 <- as.character(dfGO$Entrezid[which(dfGO$group == 3)])
selectedEntrezIds_clust4 <- as.character(dfGO$Entrezid[which(dfGO$group == 4)])

#setting cutoff value
hgCutoff <- 0.001
#hgCutoff <- 0.1

#selecting onthology

ontology <- 'BP'
#ontology <- 'MF'

params_clust2 <- new("GOHyperGParams",
 geneIds=selectedEntrezIds_clust2,
 universeGeneIds=entrezUniverse,
 annotation="org.EcK12.eg.db",
 ontology=ontology,
 pvalueCutoff=hgCutoff,
 conditional=FALSE,
 testDirection="over")

params_clust3 <- new("GOHyperGParams",
 geneIds=selectedEntrezIds_clust3,
 universeGeneIds=entrezUniverse,
 annotation="org.EcK12.eg.db",
 ontology=ontology,
 pvalueCutoff=hgCutoff,
 conditional=FALSE,
 testDirection="over")

params_clust4 <- new("GOHyperGParams",
 geneIds=selectedEntrezIds_clust4,
 universeGeneIds=entrezUniverse,
 annotation="org.EcK12.eg.db",
 ontology=ontology,
 pvalueCutoff=hgCutoff,
 conditional=FALSE,
 testDirection="over")

#We would also like to perform a conditional test.  Instead of having to de ne a new
#GO-HyperGParamsinstance  by  hand,  we  can  create  a  copy  and  update  just  the  parameter  of interest.
#paramsCond <- params
#conditional(paramsCond) <- TRUE
#hgCondOver <- hyperGTest(paramsCond)



hgOver_clust2 <- hyperGTest(params_clust2)
hgOver_clust3 <- hyperGTest(params_clust3)
hgOver_clust4 <- hyperGTest(params_clust4)



df_clust2 <- summary(hgOver_clust2)
df_clust3 <- summary(hgOver_clust3)
df_clust4 <- summary(hgOver_clust4)

########cbind(rbind(df_clust2, df_clust3, df_clust4), (df_clust2, df_clust3, df_clust4)

names(df_clust2)
#View(summary(hgOver, pvalue=0.01))
```

##Clusterization with large cluster excluded
```{r E. coli sidd, warning=FALSE, message=FALSE, cache=TRUE}
# sidds_syn <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/sidds_synechocystis_1500nts.csv' )[,-1]
matplot(sidds_ecoli,ty='l',lty =1, col=1, lwd =0.5)

sidds_ecoli_1500 <- sidds_ecoli[750:2250,-1]#row 1 nreeds to be removed since it is a rownumbers

ward_ecoli_1500sidd<- hclust.vector(t(sidds_ecoli_1500), method = 'ward') 
#plot(ward_ecoli_1500sidd, labels = F)
cut_ecoli_1500sidd <- cutree(ward_ecoli_1500sidd, k = 4)
#        table(cut_ecoli_1500sidd)
#removing  prmoters of cluster 2 (largest)

clusters_2_3_4_numbers <- which(cut_ecoli_1500sidd!=1)
#str(clusters_2_3_4_numbers)

# 3 small clusters plotted together

matplot(sidds_ecoli_1500[,clusters_2_3_4_numbers],ty='l',lty =1, col=1, lwd =0.5)

#clusterization of the 3 clusters profiles only

sidds_ecoli_small_clusters <- sidds_ecoli_1500[,clusters_2_3_4_numbers]
ward_ecoli_small_clusters <- hclust.vector(t(sidds_ecoli_1500[,clusters_2_3_4_numbers]), method = 'ward')

#plot(ward_ecoli_small_clusters, labels = F)


#cur and plot the new clusterization results
cuplcl(data = t(sidds_ecoli_small_clusters), method = 'ward', k = 3, main = 'E. coli SIDD profiles: 3 small clusters')
#`Sequence (nts)` <- -750:750
#par(mfrow = c(3,2))
#plot(ward_synechocystis_1500sidd, labels = F)
#for (i in unique(cut_synechocystis_1500sidd)){
  #matplot(`Sequence (nts)`, sidds_syn[, which(cut_synechocystis_1500sidd==i)], type = 'l', lty = 1, col=i)
 # abline(v = -10:10*100, lty =3, col = 'darkgrey', lwd =2)
#}
```


```{r Synechocystis 1500 nts long sidd1, warning=FALSE, message=FALSE, cache=TRUE}
# sidds_syn <- read.csv('https://raw.githubusercontent.com/FVortex/different_bacteria_physics_and_text/master/sidds_synechocystis_1500nts.csv' )[,-1]

matplot(sidds_syn,ty='l',lty =1, col=1, lwd =0.5)

ward_synechocystis_1500sidd<- hclust.vector(t(sidds_syn), method = 'ward')
#plot(ward_synechocystis_1500sidd, labels = F)
cut_synechocystis_1500sidd <- cutree(ward_synechocystis_1500sidd, k = 4)


#`Sequence (nts)` <- -750:750
#par(mfrow = c(3,2))
#plot(ward_synechocystis_1500sidd, labels = F)
#for (i in unique(cut_synechocystis_1500sidd)){
  #matplot(`Sequence (nts)`, sidds_syn[, which(cut_synechocystis_1500sidd==i)], type = 'l', lty = 1, col=i)
 # abline(v = -10:10*100, lty =3, col = 'darkgrey', lwd =2)
#}


sidds_synechocystis_1500 <- sidds_syn[750:2250,-1]#row 1 nreeds to be removed since it is a rownumbers

ward_sidds_synechocystis_1500 <- hclust.vector(t(sidds_synechocystis_1500), method = 'ward') 
#plot(ward_sidds_synechocystis_1500, labels = F)
cuplcl(data = t(sidds_synechocystis_1500), method = 'ward', k = 4, main = 'Synechocystis sp., SIDD')


cut_sidds_synechocystis_1500 <- cutree(ward_sidds_synechocystis_1500, k = 4)
#table(cut_sidds_synechocystis_1500)
#removing  prmoters of cluster 2 (largest)

synechocystis_clusters_2_3_4_numbers <- which(cut_sidds_synechocystis_1500!=1)
#str(synechocystis_clusters_2_3_4_numbers)

# 3 small clusters plotted together

par(mfrow = c(1,1),
    oma = rep(1.5,4))
`Sequence (nts)` <- -750:750
matplot(`Sequence (nts)`, sidds_synechocystis_1500[,synechocystis_clusters_2_3_4_numbers],ty='l',lty =1, col=1, lwd =0.5, main = 'Synechocystis sp., 3 small clusters SIDD\n (66 / 634) promoters', ylab =  'Opening probability')

#clusterization of the 3 clusters profiles only

sidds_synechocystis_small_clusters <- sidds_synechocystis_1500[,synechocystis_clusters_2_3_4_numbers]
ward_sidds_synechocystis_small_clusters <- hclust.vector(t(sidds_synechocystis_1500[,synechocystis_clusters_2_3_4_numbers]), method = 'ward')

#plot(ward_sidds_synechocystis_small_clusters, labels = F)


#cur and plot the new clusterization results
cuplcl(data = t(sidds_synechocystis_1500[,synechocystis_clusters_2_3_4_numbers]), method = 'ward', k = 3, main = 'Synechocystis sp., 3 small clusters SIDD')

#sums by rows for reduced (3 small clusters only) dataset


`Sequence (nts)` <- -750:750

sidds_sums_on_positions_synechocystis_3_clusters <- rowSums(sidds_synechocystis_1500[,synechocystis_clusters_2_3_4_numbers])
matplot(`Sequence (nts)`, sidds_sums_on_positions_synechocystis_3_clusters ,ty='l',lty =1, col=1, lwd = 1, main = 'Synechocystis sp., 3 small clusters SIDD\n (66 / 634) promoters', ylab =  'Opening probability')

library(diptest)
diptest::dip.test(x = sidds_sums_on_positions_synechocystis_3_clusters, simulate.p.value = T, B = 10000)
#and for complete set of Synechocystis profiles

sidds_sums_on_positions_synechocystis_all_clusters <- rowSums(sidds_synechocystis_1500)
matplot(`Sequence (nts)`, sidds_sums_on_positions_synechocystis_all_clusters ,ty='l',lty =1, col=1, lwd = 1, main = 'Synechocystis sp., 3 small clusters SIDD\n (66 / 634) promoters', ylab =  'Opening probability')

```

```{r Mycoplasma small peculiar cluster sequences, message=FALSE, fig.width = 10, fig.asp = 1}

mycoplasma_proms_strings <- sapply(mycoplasma_seqs,  function(x) {paste0(x, collapse = '')})
mycoplasma_proms_DNAStringSet <- DNAStringSet(mycoplasma_proms_strings)

mycoplasma_proms_stringdist<- stringDist(mycoplasma_proms_DNAStringSet, method = 'levenshtein')

ward_mycoplasma_proms_stringdist <- hclust(mycoplasma_proms_stringdist, method = 'ward.D2')

plot(ward_mycoplasma_proms_stringdist, labels = F) #k =6
ward_mycoplasma_proms_stringdist %>% as.dendrogram %>% color_branches(k = 6) %>% set('labels_cex', 1E-5) %>% raise.dendrogram(max(ward_mycoplasma_proms_stringdist$height)*0.2) -> to_plot_ward_mycoplasma_proms_stringdist

cut_ward_mycoplasma_proms_stringdist <- cutree(ward_mycoplasma_proms_stringdist, k = 3)
table(cut_ward_mycoplasma_proms_stringdist)


#for the strain used in Fisunov et al. no oriC coordinates available
#other strains comprise  rep_origin 1..1876 /note="putative oriC"
#the sequence appears to be present in the genome used at the same position

plot(seq_along(mycoplasma_CP006916.3_char), type = 'n', ylim = 0:1, main ='Mycoplasma gallisepticum levenstien distance, 2 small clusters (both 13 members) positions')
abline(v = 1:1876, lty =2)
abline(v = mycoplasma_proms$TSS[which(cut_ward_mycoplasma_proms_stringdist == 2)], col = 'darkred')
abline(v = mycoplasma_proms$TSS[which(cut_ward_mycoplasma_proms_stringdist == 3)], col = 'darkblue')

#both clusters fall into 2 two groups. What is the distance between their location?

range(mycoplasma_proms$Position[which(cut_ward_mycoplasma_proms_stringdist == 2)])
range(mycoplasma_proms$Position[which(cut_ward_mycoplasma_proms_stringdist == 3)])
#logos for [-50;50] interval
    tmps <- list()

for (i in unique(cut_ward_mycoplasma_proms_stringdist)){
    inds <- which(cut_ward_mycoplasma_proms_stringdist == i)
    tmp <- sapply(mycoplasma_seqs[inds], function(x) {substr(x, start = 1, stop = 400)})
    tmps[[i]] <-  tmp
}

ggseqlogo(tmps,  ncol = 1)+ scale_y_continuous(name="Информационнное содержание (бит)")+scale_x_discrete(name ="Последовательность относительно ТСТ (п.н.)", limits=1:400, labels = -200:200)+coord_fixed(ratio = 6)
```

```{r appendix:  coli O127:H6 isolate EPEC1 genome assembly, chromosome: 1 GenBank: LT903847.1 , eval=FALSE}

ecoli_LT903847.1 <- read.GenBank(access.nb = 'LT903847.1', as.character = T)[[1]]
#rep_origin      complement(741586..741817)
rep_origin <- 741586:741817

rep_origin_center <- round(mean(rep_origin))
write.fasta(paste0(ecoli_LT903847.1[(rep_origin_center-2500):(rep_origin_center+2500)], collapse = ''), file.out = '/home/jane/Документы/Misha/mol_a_2018/ecoli_LT903847.1_rep_origin_5000nts.fasta', names = F)

compet_result <- read.table('/home/jane/Документы/Misha/mol_a_2018/ecoli_LT903847.1_rep_origin_5000nts_competition_results.txt', skip = 24)

str(compet_result)


par(mfrow = c (3,1))
#plot(compet_result$V2, type = 'l')
plot(compet_result$V3, type = 'l',ylim = 0:1)
plot(compet_result$V4, type = 'l',ylim = 0:1)
plot(compet_result$V5, type = 'l',ylim = 0:1)

#and for 2000 nts-long region


rep_origin_center <- round(mean(rep_origin))
write.fasta(paste0(ecoli_LT903847.1[(rep_origin_center-1000):(rep_origin_center+1000)], collapse = ''), file.out = '/home/jane/Документы/Misha/mol_a_2018/ecoli_LT903847.1_rep_origin_2000nts.fasta', names = F)

compet_result <- read.table('/home/jane/Документы/Misha/mol_a_2018/ecoli_LT903847.1_rep_origin_2000nts_competition_results.txt', skip = 24)

str(compet_result)


par(mfrow = c (3,1))
#plot(compet_result$V2, type = 'l')
plot(compet_result$V3, type = 'l',ylim = 0:1)
plot(compet_result$V4, type = 'l',ylim = 0:1)
plot(compet_result$V5, type = 'l',ylim = 0:1)
```


```{r spline_dataset_pro.rdata to .scv}


dataset_pro_df <- as.data.frame(cbind(as.numeric(sapply(dataset_pro, FUN = function(x) {x$tss})),
as.character(sapply(dataset_pro, FUN = function(x) {x$strand})),
as.character(sapply(dataset_pro, FUN = function(x) {x$evidence})),
as.character(sapply(dataset_pro, FUN = function(x) {x$genename}))))

rownames(dataset_pro_df) <- names(dataset_pro)
colnames(dataset_pro_df) <- c('TSS', 'strand', 'evidence', 'genename')
write.csv(dataset_pro_df, file = '/home/jane/Документы/Misha/mol_a_2018/new_different_bacteria_physics_an_text /dataset_pro_df.csv')
```

